---
title: "dompattern by perturbation type"
author: "Annemarie van der Marel"
date: "2022-12-23"
output: html_document
---

We disentangle the effect of the different perturbation types, i.e., removals and reintroductions, on dominance pattern dynamics for all the social manipulation trials together (n = 15 trials). 
2021: 3 top-ranked trials
2022: 6 top-ranked
2022: 3 middle/low-ranked trials

We randomized the order of the dominance patterns across all trials for 1000 iterations. We then create a reference model of the transition of the dominance pattern before vs after removal and before vs after reintroduction. 

# load library
```{r echo=FALSE}
library(tidyverse)
library(diagram)
library(ggridges)
```



# import data
```{r }
dompattern21 <-  read.csv("output/dompatterns_21exp.csv") %>%
  rename(group=year,
         dompattern=dominance_pattern) %>%
  mutate(focalrank="top-ranked",
         obspattern = factor(dompattern, 
            levels =c("bullying", "close competitor","downward heuristic"),
            labels =c("B", "CC","DH"))) %>%
  dplyr::select(-X)
  # mutate(observed_pattern = factor(dominance_pattern, 
  #                            levels =c(  "bullying", "close competitor","downward heuristic")),
          #unique(dominance_pattern))

dompattern22 <- read.csv("output/dompatterns2022.csv") 


```
## Helper functions & data

```{r echo=FALSE}
# include calibri as text font
windowsFonts(Calibri = windowsFont("Calibri"))
#windowsFonts()


#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}



```

# removal pattern changes

## observed

2021
```{r}

# observed pattern change
obs_pattern_change <- dompattern21 %>%
  #dplyr::select( bin,  dompattern) %>%
  mutate(obs_after=lead(obspattern),
         obs_pattern_change = paste(obspattern, obs_after, sep="-")) %>%   
  rename(before=obspattern,
         after=obs_after, 
         change=obs_pattern_change) 

## removal
binchange_removal <- c(11, 15, 19)

obs_pattern_change_removal <- obs_pattern_change %>%
  filter(bin %in% binchange_removal) %>%
   group_by( group, before,after,change) %>% #group_by(group) 
  summarize(obs_n=length(change), obs_prop = obs_n/3) %>% 
  #dplyr::select(-obs_n) %>% 
  mutate(perturbation="removal",
         same=if_else(before==after, "yes", "no"))



```


2022
```{r}
dompattern22 <- dompattern22 %>% 
  mutate(focalrank=if_else(dompattern_focal=="middle-ranked", "middle/low", 
                           if_else(dompattern_focal=="bottom-ranked", "middle/low", "top-ranked")),
         obspattern = factor(dompattern, 
            levels =c("bullying", "close competitor","downward heuristic"),
            labels =c("B", "CC","DH"))) %>%
   dplyr::select(-X, -dompattern_focal) 


# observed pattern change

obs_pattern_change22 <- dompattern22 %>%
  #dplyr::select(group, bin, focalrank, obspattern) %>% #group, 
  group_by(group) %>% #group_by(group) 
  #filter(!bin %in% c(25)) %>%
  mutate(obs_after=lead(obspattern),
         obs_pattern_change = paste(obspattern, obs_after, sep="-")) %>%   
  rename(before=obspattern,
         after=obs_after, 
         change=obs_pattern_change) 


## removal
binchange_removal22 <- c(2,6,10,14, 18, 22)

obs_pattern_change_removal22 <- obs_pattern_change22 %>%
  filter(bin %in% binchange_removal22) %>%
  group_by(group, before,after,change) %>% #group_by(group) 
  summarize(obs_n=length(change), obs_prop = obs_n/6) %>% 
  #dplyr::select(-obs_n) %>% 
  mutate(perturbation="removal",
         same=if_else(before==after, "yes", "no"))

check <- obs_pattern_change_removal22 %>% 
  #group_by(focalrank) %>%
  summarise(sum(obs_prop)) # should be 1
sum(obs_pattern_change_removal$obs_n)



```
combine
```{r}
removal_change <- bind_rows(obs_pattern_change_removal, 
                            obs_pattern_change_removal22) %>%
  group_by(before, after, change) %>%
  summarize(n=sum(obs_n), obs_prop=n/15) %>%  
  mutate(prop=round(obs_prop, 2))

sum(removal_change$n)
```

### visualise
plotmat visualisez matrix (rows=to, cols=from))
I have to change the direction of my matrix as I had rows=from/before and cols=to/after. 
```{r}
## observed : NB. (rows=to, cols=from) in plotmat 
mx.removal <- reshape2::dcast(removal_change , after~before, value.var="prop") 
mx.removal <- matrix.please(mx.removal)
mx.removal[is.na(mx.removal)] <- 0

 # set line width
linewidth <- matrix(nrow=3,
                    ncol=3,0)
linewidth[1,1]<- (mx.removal[1,1]*10)+1      
linewidth[1,2]<- (mx.removal[1,2]*10)+1
linewidth[1,3]<- (mx.removal[1,3]*10)+1 
linewidth[2,2]<- (mx.removal[2,2]*10)+1  
linewidth[2,1]<- (mx.removal[2,1]*10)
linewidth[2,3]<- (mx.removal[2,3]*10)
linewidth[3,1]<- (mx.removal[3,1]*10)+1       
linewidth[3,2]<- (mx.removal[3,2]*10)+1 
linewidth[3,3]<- (mx.removal[3,3]*10)+1 



# plot transition matrix
pdf("./figures/plot_part4_obs_removal.pdf")  # open and save the pdf file
plotmat(mx.removal, 
        #pty="s",
        #pos=c(1,1),
        
        # lines
        box.lwd=1, lwd=linewidth, lcol = "#ec0000bf",  #lcol=c( "#0000FF", "#7F7F7F"), 
        
        # boxes
        box.cex=0.8, box.size=0.065 ,box.type="rect", # "circle", "ellipse", "diamond", "rect", "round", "hexa", "multi", "none"
        box.prop=1, box.col = c( "#0000FF","#C00000" , "#7F7F7F"),
        shadow.size = 0.0, 
        my=0, mx=0.0, # vertical and horizontal shift of boxes, respecitvely 
        
        # self 
        relsize=0.7, self.lwd=c(4.3,4.3), #self.lwd = linewidth, 
        self.shiftx=c(.1, -0.1), self.shifty=c(-0.1, 0.1), 
        
        # arrow
        #arr.lwd = c(1, 2,3),
        #arr.length = 2,
        curve = 0.125,
        arr.tcol = "black" , arr.col = "#ec0000bf",
     
        # text
        cex.txt=0.8, txt.col = "white",
        main = "observed transitions after removals"
        )
dev.off() # Close the pdf file



```







# reintroduction pattern change

## observed
```{r}
## reintroduction 2021
obs_pattern_change <- dompattern21 %>%
  #dplyr::select( bin,  dompattern) %>%
  mutate(obs_after=lead(obspattern),
         obs_pattern_change = paste(obspattern, obs_after, sep="-")) %>%   
  rename(before=obspattern,
         after=obs_after, 
         change=obs_pattern_change) 

binchange_reintro <- c(13, 17, 21)

obs_pattern_change_reintro <- obs_pattern_change %>%
  filter(bin %in% binchange_reintro) %>%
   group_by( group, before,after,change) %>% #group_by(group) 
  summarize(obs_n=length(change), obs_prop = obs_n/3) %>% 
  #dplyr::select(-obs_n) %>% 
  mutate(perturbation="removal",
         same=if_else(before==after, "yes", "no"))

## reintroduction 2022
binchange_reintro22 <- c(4, 8, 12, 16, 20, 24)

obs_pattern_change_reintro22 <- obs_pattern_change22 %>%
  filter(bin %in% binchange_reintro22) %>%
  group_by(focalrank, before,after,change) %>% #group_by(group) 
  summarize(obs_n=length(change), obs_prop = obs_n/6) %>% 
  #dplyr::select(-obs_n) %>% 
  mutate(perturbation="reintroduction",
         same=if_else(before==after, "yes", "no")) 

sum(obs_pattern_change_reintro22$obs_n)



```

combine
```{r}
reintro_change <- bind_rows(obs_pattern_change_reintro, 
                            obs_pattern_change_reintro22) %>%
  group_by(before, after, change) %>%
  summarize(n=sum(obs_n), obs_prop=n/15) %>%
  mutate(prop=round(obs_prop, 2))

sum(reintro_change$n)
```


### visualise
plotmat visualisez matrix (rows=to, cols=from))
I have to change the direction of my matrix as I had rows=from/before and cols=to/after. 
```{r}
## observed : NB. (rows=to, cols=from) in plotmat 
mx.reintro <- reshape2::dcast(reintro_change , after~before, value.var="prop") 
mx.reintro <- matrix.please(mx.reintro)
mx.reintro[is.na(mx.reintro)] <- 0

 # set line width
linewidth <- matrix(nrow=3,
                    ncol=3,0)
linewidth[1,1]<- (mx.reintro[1,1]*10)+1      
linewidth[1,2]<- (mx.reintro[1,2]*10)+1 
linewidth[1,3]<- (mx.reintro[1,3]*10)+1 
linewidth[2,2]<- (mx.reintro[2,2]*10)+1  
linewidth[2,1]<- (mx.reintro[2,1]*10)
linewidth[2,3]<- (mx.reintro[2,3]*10)  
linewidth[3,1]<- (mx.reintro[3,1]*10)+1       
linewidth[3,2]<- (mx.reintro[3,2]*10)+1
linewidth[3,3]<- (mx.reintro[3,3]*10)+1 



# plot transition matrix
pdf("./figures/plot_part4_obs_reintro.pdf")  # open and save the pdf file
plotmat(mx.reintro, 
        #pty="s",
        #pos=c(1,1),
        
        # lines
        box.lwd=1, lwd=linewidth, lcol = "#ec0000bf",  #lcol=c( "#0000FF", "#7F7F7F"), 
        
        # boxes
        box.cex=0.8, box.size=0.065 ,box.type="rect", # "circle", "ellipse", "diamond", "rect", "round", "hexa", "multi", "none"
        box.prop=1, box.col = c( "#0000FF","#C00000" , "#7F7F7F"),
        shadow.size = 0.0, 
        my=0, mx=0.0, # vertical and horizontal shift of boxes, respecitvely 
        
        # self 
        relsize=0.7, self.lwd=c(4.3,4.3), #self.lwd = linewidth, 
        self.shiftx=c(.1, -0.1), self.shifty=c(-0.1, 0.1), 
        
        # arrow
        #arr.lwd = c(1, 2,3),
        #arr.length = 2,
        curve = 0.125,
        arr.tcol = "black" , arr.col = "#ec0000bf",
     
        # text
        cex.txt=0.8, txt.col = "white",
        main = "observed transitions after reintroductions"
        )
dev.off() # Close the pdf file



```

# Randomise pattern changes

Permute social dominance pattern points

-   Change order of the data points by group

## 2021
```{r}
# randomise dom patterns -----

glimpse(dompattern21)

# Step 1: create new df
s.pattern <- dompattern21 %>%
  dplyr::select(group, bin, obspattern, focalrank) %>%
  rename(observed_pattern=obspattern)


# Step 2: create loop to get 1000 randomised patterns
random_patterns <- data.frame(runID =numeric(),
                              group = numeric(),
                              focalrank = character(),
                              bin=numeric(),
                              oberved_pattern=character(),
                              random_pattern=character())

replicates <- 1000

for(run in 1:replicates){
 r.seed <- run
 set.seed(r.seed)

# create random distribution dominance patterns
  s.pattern$runID <- rep(paste0("run",str_pad(r.seed, 4, side="left", pad = "0")))
	
  samplepat <- sample(1:length(s.pattern$observed_pattern))
	samplepat
	s.pattern$random_pattern <- s.pattern$observed_pattern[samplepat]

  random_patterns <- rbind.data.frame(random_patterns, s.pattern)
  
}


## check ref model output

#check that the randomized interactions vary from the observed patterns
random_patterns$observed_pattern == random_patterns$random_pattern

#check that runs are different from each other
head(subset(random_patterns, runID=="run0001"))
head(subset(random_patterns, runID=="run0002"))
head(subset(random_patterns, runID=="run0003"))
head(subset(random_patterns, runID=="run0011"))
head(subset(random_patterns, runID=="run0100"))

#check that all runs ran as expected
unique(random_patterns$runID)
length(unique(random_patterns$runID))


# Step 3: Calculate how often we see the same strategies or a change to bullying 

before_after21 <- random_patterns %>% 
  dplyr::select(group, runID, bin, focalrank, observed_pattern, random_pattern) %>%

  # observed pattern change 
  rename(obs_before = observed_pattern) %>%
  group_by(runID) %>%
  mutate(obs_after = lead(obs_before)) %>%
  #drop_na() %>%
  mutate(obs_pattern_change = paste(obs_before, obs_after, sep="-")) %>%

# randomized pattern change
  rename(random_before = random_pattern) %>%
  group_by(runID) %>%
  mutate(random_after = lead(random_before)) %>%
  mutate(random_pattern_change = paste(random_before, random_after, sep="-")) %>%
  drop_na() %>%
  
  
  dplyr:: select(group, runID, bin, focalrank, obs_pattern_change,  random_pattern_change)

#write.csv(before_after, "output/dompattern_random.csv")

glimpse(before_after21)

```

### removal
```{r}
glimpse(before_after21)

## ----

# Step 2: create loop to get summary
rand_summ <- data.frame(
                              group = numeric(),
                              focalrank = character(),
                              bin=numeric(),
                              change=character(),
                              random_n=numeric(),
                              prop=numeric())


## filter by bin/trial
binchange_removal <- c(11, 15, 19)

for(i in seq_along(binchange_removal)){
  runby <- binchange_removal[i]
  print(runby)
  run.data <- subset(before_after21, bin==runby)
  
  
  ## summarize
pool<- run.data %>%
  group_by(group, bin, random_pattern_change) %>%
  summarise(random_n=length(random_pattern_change), 
            random_prop = random_n/1000) %>%  # each run with 1000 randomized values
  rename(change=random_pattern_change) %>%
  #dplyr::select(-random_n) %>%
  mutate(prop=round(random_prop, 2)
         ) %>%
  dplyr::select(-random_prop)

check<- pool %>%
  summarize(sum=sum(random_n))
print(check$sum) # should be 1000

 rand_summ<-rbind(rand_summ, pool )


}

rand_summ21 <- rand_summ

```

### reintroduction

```{r}

glimpse(before_after21)

## ----

# Step 2: create loop to get summary
rand_summ <- data.frame(
                              group = numeric(),
                              focalrank = character(),
                              bin=numeric(),
                              change=character(),
                              random_n=numeric(),
                              prop=numeric())


## filter by bin/trial
binchange_reintro <- c(13, 17, 21)

for(i in seq_along(binchange_reintro)){
  runby <- binchange_reintro[i]
  print(runby)
  run.data <- subset(before_after21, bin==runby)
  
  
  ## summarize
pool<- run.data %>%
  group_by( group, bin, random_pattern_change) %>%
  summarise(random_n=length(random_pattern_change), 
            random_prop = random_n/1000) %>%  # each run with 1000 randomized values
  rename(change=random_pattern_change) %>%
  #dplyr::select(-random_n) %>%
  mutate(prop=round(random_prop, 2)
         ) %>%
  dplyr::select(-random_prop)

check<- pool %>%
  summarize(sum=sum(random_n))
print(check$sum) # should be 1000

 rand_summ<-rbind(rand_summ, pool )


}

rand_reintro21 <- rand_summ
```


## 2022

```{r randomise strategies 2022}
# dompattern22 <- dompattern22 %>% 
#   mutate(focalrank=if_else(dompattern_focal=="middle-ranked", "middle/low", 
#                            if_else(dompattern_focal=="bottom-ranked", "middle/low", "top-ranked")),
#          obspattern = factor(dompattern, 
#             levels =c("bullying", "close competitor","downward heuristic"),
#             labels =c("B", "CC","DH"))) %>%
#    dplyr::select(-X, -dompattern_focal) 

# randomise dom patterns -----
# Step 1: create new df
s.pattern <- dompattern22 %>%
  dplyr::select(group, bin, obspattern, focalrank) %>%
  filter(!bin %in% c(1,26)) %>%
  rename(observed_pattern=obspattern)


# Step 2: create loop to get 1000 randomised patterns
random_patterns1 <- data.frame(runID =numeric(),
                              group = numeric(),
                              focalrank = character(),
                              bin=numeric(),
                              oberved_pattern=character(),
                              random_pattern=character())


# by group separately
## group1
run.data1 <- subset(s.pattern, group==1)

replicates <- 1000

for(run in 1:replicates){
 r.seed <- run
 set.seed(r.seed)

# create random distribution dominance patterns
   run.data1$runID <- rep(paste0("run",str_pad(r.seed, 4, side="left", pad = "0")))
	
  samplepat <- sample(1:length( run.data1$observed_pattern))
	samplepat
	 run.data1$random_pattern <-  run.data1$observed_pattern[samplepat]

  random_patterns1 <- rbind.data.frame(random_patterns1,  run.data1)
  
}

## group2
random_patterns2 <- data.frame(runID =numeric(),
                              group = numeric(),
                              focalrank = character(),
                              bin=numeric(),
                              oberved_pattern=character(),
                              random_pattern=character())

run.data2 <- subset(s.pattern, group==2)

replicates <- 1000

for(run in 1:replicates){
 r.seed <- run
 set.seed(r.seed)

# create random distribution dominance patterns
   run.data2$runID <- rep(paste0("run",str_pad(r.seed, 4, side="left", pad = "0")))
	
  samplepat <- sample(1:length( run.data2$observed_pattern))
	samplepat
	 run.data2$random_pattern <-  run.data2$observed_pattern[samplepat]

  random_patterns2 <- rbind.data.frame(random_patterns2,  run.data2)
  
}

## combine
random_patterns <- bind_rows(random_patterns1,
                             random_patterns2)


## check ref model output


# check whether per run and per group it adds up to the same number of patterns
check_random1 <- random_patterns1 %>%
  group_by(runID, random_pattern) %>%
  summarize(n=length(random_pattern)) %>% # somehow doubling the totals
  ungroup()

# check whether per run and per group it adds up to the same number of patterns
check_random2 <- random_patterns2 %>%
  group_by(runID, random_pattern) %>%
  summarize(n=length(random_pattern)) %>% # somehow doubling the totals
  ungroup()

# observed patterns per group
dompattern_summ <- dompattern22 %>%
  group_by(group, dompattern) %>%
  filter(!bin %in% c(1,26)) %>%
  tally()

  

#check that the randomized interactions vary from the observed patterns
random_patterns$observed_pattern == random_patterns$random_pattern

#check that runs are different from each other
head(subset(random_patterns, runID=="run0001"))
head(subset(random_patterns, runID=="run0002"))
head(subset(random_patterns, runID=="run0003"))
head(subset(random_patterns, runID=="run0011"))
head(subset(random_patterns, runID=="run0100"))

#check that all runs ran as expected
unique(random_patterns$runID)
length(unique(random_patterns$runID))


# Step 3: Calculate how often do we see the same strategies or a change to bullying 

before_after22 <- random_patterns %>% 
  dplyr::select(runID, group, bin, focalrank, observed_pattern, random_pattern) %>%

  # observed pattern change 
  rename(obs_before = observed_pattern) %>%
  group_by(runID, group) %>%
  mutate(obs_after = lead(obs_before)) %>%
  #drop_na() %>%
  mutate(obs_pattern_change = paste(obs_before, obs_after, sep="-")) %>%

# randomized pattern change
  rename(random_before = random_pattern) %>%
  group_by(runID, group) %>%
  mutate(random_after = lead(random_before)) %>%
  mutate(random_pattern_change = paste(random_before, random_after, sep="-")) %>%
  drop_na() %>%
  
  
  dplyr:: select(group, runID,bin, focalrank,  obs_pattern_change,  random_pattern_change)

#write.csv(before_after, "output/dompattern_random.csv")

glimpse(before_after22)

```

### removal

```{r }

glimpse(before_after22)

## ----

# Step 2: create loop to get summary
rand_summ <- data.frame(
                              group = numeric(),
                              focalrank = character(),
                              bin=numeric(),
                              change=character(),
                              random_n=numeric(),
                              prop=numeric())


## filter by group
bygroup <- unique(before_after22$group)

for(i in seq_along(bygroup)){
  run.code <- bygroup[i]
  print(run.code)
  group.data <- subset(before_after22, group==run.code)

## filter by bin/trial
binchange_removal22 <- c(2,6,10,14, 18, 22)

for(i in seq_along(binchange_removal22)){
  runby <- binchange_removal22[i]
  print(runby)
  run.data <- subset(group.data, bin==runby)
  
  
  ## summarize
pool<- run.data %>%
  group_by(group, bin, random_pattern_change) %>%
  summarise(random_n=length(random_pattern_change), 
            random_prop = random_n/1000) %>%  # each run with 1000 randomized values
  rename(change=random_pattern_change) %>%
  #dplyr::select(-random_n) %>%
  mutate(prop=round(random_prop, 2)
         ) %>%
  dplyr::select(-random_prop)

check<- pool %>%
  summarize(sum=sum(random_n))
print(check$sum) # should be 1000

 rand_summ<-rbind(rand_summ, pool )


}}

rand_summ22 <- rand_summ

```





### reintroduction


```{r }

glimpse(before_after22)

## ----

# Step 2: create loop to get summary
rand_summ <- data.frame(
                              group = numeric(),
                              focalrank = character(),
                              bin=numeric(),
                              change=character(),
                              random_n=numeric(),
                              prop=numeric())


## filter by group
bygroup <- unique(before_after22$group)

for(i in seq_along(bygroup)){
  run.code <- bygroup[i]
  print(run.code)
  group.data <- subset(before_after22, group==run.code)

## filter by bin/trial
binchange_reintro22 <- c(4, 8, 12, 16, 20, 24)

for(i in seq_along(binchange_reintro22)){
  runby <- binchange_reintro22[i]
  print(runby)
  run.data <- subset(group.data, bin==runby)
  
  
  ## summarize
pool<- run.data %>%
  group_by(group, bin, random_pattern_change) %>%
  summarise(random_n=length(random_pattern_change), 
            random_prop = random_n/1000) %>%  # each run with 1000 randomized values
  rename(change=random_pattern_change) %>%
  #dplyr::select(-random_n) %>%
  mutate(prop=round(random_prop, 2)
         ) %>%
  dplyr::select(-random_prop)

check<- pool %>%
  summarize(sum=sum(random_n))
print(check$sum) # should be 1000

 rand_summ<-rbind(rand_summ, pool )


}}

rand_reintro22 <- rand_summ

```


# Plots
## removal

###Combine
```{r}
rand_summ_removal <- bind_rows(rand_summ21, 
                               rand_summ22)

rand_removal <- rand_summ_removal %>%
   separate(change, into = c("before", "after"), sep="-", remove = F)  %>%
  group_by(before,after, change) %>% 
 
  summarize(aveprop=mean(prop),
            sdprop=sd(prop),
            n=sum(random_n),
            se=sdprop/sqrt(n)) %>%
  mutate(prop=round(aveprop, 2),
         sd=round(sdprop, 2))

sum(rand_removal$aveprop)
```

###transition diagram

```{r}

## observed : NB. (rows=to, cols=from) in plotmat 
mx.removal.rand <- reshape2::dcast(rand_removal , after~before, value.var="prop") 
mx.removal.rand <- matrix.please(mx.removal.rand )
mx.removal.rand [is.na(mx.removal.rand )] <- 0

 # set line width
linewidth <- matrix(nrow=3,
                    ncol=3,0)
linewidth[1,1]<- (mx.removal.rand [1,1]*10)+1      
linewidth[1,2]<- (mx.removal.rand [1,2]*10)+1 
linewidth[1,3]<- (mx.removal.rand [1,3]*10)+1 
linewidth[2,2]<- (mx.removal.rand [2,2]*10)+1  
linewidth[2,1]<- (mx.removal.rand [2,1]*10)+1
linewidth[2,3]<- (mx.removal.rand [2,3]*10)+1 
linewidth[3,1]<- (mx.removal.rand [3,1]*10)+1       
linewidth[3,2]<- (mx.removal.rand [3,2]*10)+1
linewidth[3,3]<- (mx.removal.rand [3,3]*10)+1 



# plot transition matrix
pdf("./figures/plot_part4_rand_removal.pdf")  # open and save the pdf file
plotmat(mx.removal.rand , 
        #pty="s",
        #pos=c(1,1),
        
        # lines
        box.lwd=1, lwd=linewidth, lcol = "#ec0000bf",  #lcol=c( "#0000FF", "#7F7F7F"), 
        
        # boxes
        box.cex=0.8, box.size=0.065 ,box.type="rect", # "circle", "ellipse", "diamond", "rect", "round", "hexa", "multi", "none"
        box.prop=1, box.col = c( "#0000FF","#C00000" , "#7F7F7F"),
        shadow.size = 0.0, 
        my=0, mx=0.0, # vertical and horizontal shift of boxes, respecitvely 
        
        # self 
        relsize=0.7, self.lwd=c(4.3,4.3), #self.lwd = linewidth, 
        self.shiftx=c(.1, -0.1), self.shifty=c(-0.1, 0.1), 
        
        # arrow
        #arr.lwd = c(1, 2,3),
        #arr.length = 2,
        curve = 0.125,
        arr.tcol = "black" , arr.col = "#ec0000bf",
     
        # text
        cex.txt=0.8, txt.col = "white",
        main = "randomised transitions after removals"
        )
dev.off() # Close the pdf file



```


### distribution
Double-check analysis -> now separate per group but should I not include all in one and than calculate proportions so there is not a 3-model distrition but a unimodal distribution. So when I randomise per group and get the randomised pattern changes, I have to combine the datasets and divide by ... 


```{r}
 plot_dist_removal <-  ggplot(rand_summ_removal, 
                              aes(x = prop, y = change, fill = change)) +
  
  geom_density_ridges2(scale=0.8, draw_baseline = F) +
  
  theme_ridges(#font_size = 12,
    #font_family = "",
   line_size = 0.01,
   # grid = FALSE,
    center_axis_labels = FALSE) + #grid=FALSE
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      axis.title.x = element_text(hjust = 0.5)
      #axis.title.y = element_text(hjust = 0.5)
    ) +
    xlab("proportion of runs") +
    ylab("") +
    #scale_fill_manual(values = c("#0000FF", "#b3b0e3", "#deebf7", "#7F7F7F")) + #"#C00000" ,
  # scale_x_continuous( expand = c(0, 0),limits = c(-0.1,1.2)) +
   scale_y_discrete(expand = expansion(mult = c(0.0, .3))) +
   
  # include observed proportion top-ranked (color orange)
  geom_segment(y=9, yend=9.8, x = 0.27, xend=0.27, size=1) + # DH-DH 
  # B-CC NA
  geom_segment(y=7, yend=7.8, x = 0.07, xend=0.07, size=1) + # B-DH 
   geom_segment(y=6, yend=6.8, x = 0.13, xend=0.13, size=1) + # CC-B
   geom_segment(y=5, yend=5.8, x = 0.13, xend=0.13, size=1) + #CC-CC
  geom_segment(y=4, yend=4.8, x = 0.13, xend=0.13, size=1) +  # CC-DH
  geom_segment(y=3, yend=3.8, x = 0.13, xend=0.13, size=1) + # DH-B
  #Dh-CC
   geom_segment(y=1, yend=1.8, x = 0.13, xend=0.13, size=1) + # B-B
   
   
  #geom_segment(y=1, yend=5, x = 1, xend=1, size=1,  color="white") +
   coord_cartesian(clip = "off") 

plot_dist_removal


ggsave("figures/plot_dist_removal_v1.pdf", height=4, width = 4)
```


```{r}
 plot_dist_removal <-  ggplot(rand_summ_removal, 
                              aes(x = prop, y = change, fill = change)) +
  
  geom_density_ridges( stat="binline",
                       bins=8.5, scale=0.8, draw_baseline = T,  center = 0) +    #alpha=0.6, binwidth=1
  theme_ridges(#font_size = 12,
    #font_family = "",
   line_size = 0.01,
   # grid = FALSE,
    center_axis_labels = FALSE) + #grid=FALSE
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      axis.title.x = element_text(hjust = 0.5)
      #axis.title.y = element_text(hjust = 0.5)
    ) +
    xlab("proportion of runs") +
    ylab("") +
    #scale_fill_manual(values = c("#0000FF", "#b3b0e3", "#deebf7", "#7F7F7F")) + #"#C00000" ,
   scale_x_continuous( expand = c(0, 0),limits = c(-0.1,1.2),
                       breaks = c(0.00, 0.33, 0.66, 1.00)) +
   scale_y_discrete(expand = expansion(mult = c(0.0, .3))) +
   
  # include observed proportion top-ranked (color orange)
  geom_segment(y=9, yend=9.8, x = 0.27, xend=0.27, size=1) + # DH-DH 
  # B-CC NA
  geom_segment(y=7, yend=7.8, x = 0.07, xend=0.07, size=1) + # B-DH 
   geom_segment(y=6, yend=6.8, x = 0.13, xend=0.13, size=1) + # CC-B
   geom_segment(y=5, yend=5.8, x = 0.13, xend=0.13, size=1) + #CC-CC
  geom_segment(y=4, yend=4.8, x = 0.13, xend=0.13, size=1) +  # CC-DH
  geom_segment(y=3, yend=3.8, x = 0.13, xend=0.13, size=1) + # DH-B
  #Dh-CC
   geom_segment(y=1, yend=1.8, x = 0.13, xend=0.13, size=1) + # B-B
   
   
  #geom_segment(y=1, yend=5, x = 1, xend=1, size=1,  color="white") +
   coord_cartesian(clip = "off") 

plot_dist_removal


ggsave("figures/plot_dist_removal.pdf")
```



## reintroduction

### combine
```{r}

rand_summ_reintro <- bind_rows(rand_reintro21, 
                               rand_reintro22)

rand_reintro <- rand_summ_reintro %>%
     separate(change, into = c("before", "after"), sep="-", remove = F)  %>%
  group_by(before,after, change) %>% 
  
  summarize(aveprop=mean(prop),
            sdprop=sd(prop),
            n=sum(random_n),
            se=sdprop/sqrt(n)) %>%
  mutate(prop=round(aveprop, 2),
         sd=round(sdprop, 2))

sum(rand_reintro$prop)


```



### transition diagram
```{r}

## observed : NB. (rows=to, cols=from) in plotmat 
mx.reintro.rand <- reshape2::dcast(rand_reintro , after~before, value.var="prop") 
mx.reintro.rand <- matrix.please(mx.reintro.rand )
mx.reintro.rand [is.na(mx.reintro.rand )] <- 0

 # set line width
linewidth <- matrix(nrow=3,
                    ncol=3,0)
linewidth[1,1]<- (mx.reintro.rand [1,1]*10)+1      
linewidth[1,2]<- (mx.reintro.rand [1,2]*10)+1 
linewidth[1,3]<- (mx.reintro.rand [1,3]*10)+1 
linewidth[2,2]<- (mx.reintro.rand [2,2]*10)+1  
linewidth[2,1]<- (mx.reintro.rand [2,1]*10)+1
linewidth[2,3]<- (mx.reintro.rand [2,3]*10)+1 
linewidth[3,1]<- (mx.reintro.rand [3,1]*10)+1       
linewidth[3,2]<- (mx.reintro.rand [3,2]*10)+1
linewidth[3,3]<- (mx.reintro.rand [3,3]*10)+1 



# plot transition matrix
pdf("./figures/plot_part4_rand_reintro.pdf")  # open and save the pdf file
plotmat(mx.reintro.rand , 
        #pty="s",
        #pos=c(1,1),
        
        # lines
        box.lwd=1, lwd=linewidth, lcol = "#ec0000bf",  #lcol=c( "#0000FF", "#7F7F7F"), 
        
        # boxes
        box.cex=0.8, box.size=0.065 ,box.type="rect", # "circle", "ellipse", "diamond", "rect", "round", "hexa", "multi", "none"
        box.prop=1, box.col = c( "#0000FF","#C00000" , "#7F7F7F"),
        shadow.size = 0.0, 
        my=0, mx=0.0, # vertical and horizontal shift of boxes, respecitvely 
        
        # self 
        relsize=0.7, self.lwd=c(4.3,4.3), #self.lwd = linewidth, 
        self.shiftx=c(.1, -0.1), self.shifty=c(-0.1, 0.1), 
        
        # arrow
        #arr.lwd = c(1, 2,3),
        #arr.length = 2,
        curve = 0.125,
        arr.tcol = "black" , arr.col = "#ec0000bf",
     
        # text
        cex.txt=0.8, txt.col = "white",
        main = "randomised transitions after reintroductions"
        )
dev.off() # Close the pdf file



```

### distribution 

```{r}
plot_dist_reintro <-  ggplot(rand_summ_reintro, 
                              aes(x = prop, y = change, fill = change)) +
  geom_density_ridges2(scale=0.8, draw_baseline = F) +
  
  theme_ridges(#font_size = 12,
    #font_family = "",
   line_size = 0.01,
   # grid = FALSE,
    center_axis_labels = FALSE) + #grid=FALSE
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      axis.title.x = element_text(hjust = 0.5)
      #axis.title.y = element_text(hjust = 0.5)
    ) +
    xlab("proportion of runs") +
    ylab("") +
    #scale_fill_manual(values = c("#0000FF", "#b3b0e3", "#deebf7", "#7F7F7F")) + #"#C00000" ,
  # scale_x_continuous( expand = c(0, 0),limits = c(-0.1,1.2)) +
   scale_y_discrete(expand = expansion(mult = c(0.0, .3))) +
   
   # include observed proportion top-ranked (color orange)
  geom_segment(y=9, yend=9.8, x = 0.13, xend=0.13, size=1) + # DH-DH 
  # B-CC NA
  geom_segment(y=7, yend=7.8, x = 0.33, xend=0.33, size=1) + # B-DH 
   geom_segment(y=6, yend=6.8, x = 0.13, xend=0.13, size=1) + # CC-B
   geom_segment(y=5, yend=5.8, x = 0.13, xend=0.13, size=1) + #CC-CC
  geom_segment(y=4, yend=4.8, x = 0.07, xend=0.07, size=1) +  # CC-DH
  geom_segment(y=3, yend=3.8, x = 0.07, xend=0.07, size=1) + # DH-B
  #Dh-CC
   geom_segment(y=1, yend=1.8, x = 0.13, xend=0.13, size=1) + # B-B
   
  #geom_segment(y=1, yend=5, x = 1, xend=1, size=1,  color="white") +
   coord_cartesian(clip = "off") 

plot_dist_reintro

ggsave("figures/plot_dist_reintro.pdf", height = 4, width = 4)
```



```{r}
 plot_dist_reintro <-  ggplot(rand_summ_reintro, 
                              aes(x = prop, y = change, fill = change)) +
  
  geom_density_ridges( stat="binline",
                       bins=8.5, scale=0.8, draw_baseline = F, center = 0) +    #alpha=0.6, binwidth=1
  theme_ridges(#font_size = 12,
    #font_family = "",
   line_size = 0.01,
   # grid = FALSE,
    center_axis_labels = FALSE) + #grid=FALSE
  theme(
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      strip.text.x = element_text(size = 8),
      axis.title.x = element_text(hjust = 0.5)
      #axis.title.y = element_text(hjust = 0.5)
    ) +
    xlab("proportion of runs") +
    ylab("") +
    #scale_fill_manual(values = c("#0000FF", "#b3b0e3", "#deebf7", "#7F7F7F")) + #"#C00000" ,
   scale_x_continuous( expand = c(0, 0),limits = c(-0.1,1.2),
                       breaks = c(0.00, 0.33, 0.66, 1.00)) +
   scale_y_discrete(expand = expansion(mult = c(0.0, .3))) +
   
  # include observed proportion top-ranked (color orange)
  geom_segment(y=9, yend=9.8, x = 0.13, xend=0.13, size=1) + # DH-DH 
  # B-CC NA
  geom_segment(y=7, yend=7.8, x = 0.33, xend=0.33, size=1) + # B-DH 
   geom_segment(y=6, yend=6.8, x = 0.13, xend=0.13, size=1) + # CC-B
   geom_segment(y=5, yend=5.8, x = 0.13, xend=0.13, size=1) + #CC-CC
  geom_segment(y=4, yend=4.8, x = 0.07, xend=0.07, size=1) +  # CC-DH
  geom_segment(y=3, yend=3.8, x = 0.07, xend=0.07, size=1) + # DH-B
  #Dh-CC
   geom_segment(y=1, yend=1.8, x = 0.13, xend=0.13, size=1) + # B-B
   
   
  #geom_segment(y=1, yend=5, x = 1, xend=1, size=1,  color="white") +
   coord_cartesian(clip = "off") 

plot_dist_reintro



 ```

