---
title: "Aggression networks"
author: "Annemarie van der Marel"
date: "2024-12-10"
output: html_document
---

#Session info
```{r}
sessionInfo()

#citation('diagram')
#citation('ggridges')
```


#Load packages
```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(reshape2)
library(stringr)
library(purrr)
library(domstruc) #install_github("danm0nster/domstruc")
library(ggraph)
library(diagram)
library(ggpubr)
library(ggridges)
library(igraph)
library(ggraph)
library(vegan)
```

# import data

```{r}
aggXbin21 <- read.csv("./output/dyad.aggXbin.socialexp21.csv") %>%
  dplyr::select(-X)


aggXbin22 <- read.csv("./output/dyad.aggXbin2022.csv") %>%
  dplyr::select(-X)

# power scores 
power21 <- read.csv("./output/MOPA_power score_2021.csv") %>% 
  rename(period=bin)

power22 <- read.csv("./output/powerXbin2022.csv")

bins21 <- read.csv("./data/bins2021.csv")

# focals 2022
focals22 <- read.csv("./data/2022_focal_birds.csv") 

```



# 2021 Aggression networks

join dominance scores to dyad aggression dataframe for both actor and subject

Structure of hierarchy:
- % of fights rule breakers vs rule followers

plot aggression network by dominance rank on y-axis and jitter on x-axis. One convenient way to adjust node coordinates is to plot with tkplot first, handtune the placement of the vertices, query the coordinates by the tk_coords function and use them with plot to plot the graph to any R device.

## bin selection
before vs after removal
BBB: bin c(1,2)
GPG: bin c(5,6)
OPP: bin c(9,10)

before reintroduction vs upon reintroduction 
BBB: bin  c(3,4) 
GPG: bin  c(7,8)
OPP: bin  c(11,12)
```{r}
bin19birds <- c(2,3,6,7,10,11) # include removed focal bird in matrix with 0's
bin20birds <- c(1,4,5,8,9,12)

# per trial and period separate -> differently colored focals compared to the rest of the birds
bintrial1_focal <- c(1,4)
bintrial1_nofocal <- c(2,3)
bintrial2_focal <- c(5,8)
bintrial2_nofocal <- c(6,7)
bintrial3_focal <- c(9,12)
bintrial3_nofocal <- c(10,11)

```

## prepare rank dataframe
```{r}
glimpse(ec_rankXbin)
range(ec_rankXbin$dom_ec)

# change bins
bins <- bin.dates %>%
  dplyr::select(bin, rank_assessment) %>%
  drop_na() %>%
  mutate(bin=as.integer(bin)) %>%
  group_by(bin) %>%
  slice(1)

ec_rank<- ec_rankXbin %>%
  left_join(bins) %>% 
  dplyr::select(-bin) %>%
  rename(bin=rank_assessment) %>%
  mutate(power=1-dom_ec,
         bin=as.factor(bin)) %>%
  dplyr::select(bin, id,power, dom_ec, rank)
  
glimpse(ec_rank)

```

## create edge list
change bin each time as the loop didn't work
```{r}
# check data
glimpse(dyad.totalaggXbin)
sum(dyad.totalaggXbin$n.aggXbin)
unique(dyad.totalaggXbin$subject)
length(unique(dyad.totalaggXbin$actor))
length(unique(dyad.totalaggXbin$subject))
sort(unique(dyad.totalaggXbin$bin))

dyad.totalaggXbin$bin <- as.factor(dyad.totalaggXbin$bin)

# combine dom_ec and rank to edge  dataframe
attrib_edge <- dyad.totalaggXbin %>%
  dplyr::select(-X) %>%
  #filter(bin %in% bintrial2_nofocal) %>%   # change 
  filter(bin==8) %>%
  #slice(rep(seq_len(n()), n.aggXbin)) %>%  # transform to edge list
  #mutate(weight = 1) %>%
  dplyr::rename(id=actor, weight=n.aggXbin) %>% 
  left_join(ec_rank, by=c("id", "bin")) %>% 
  dplyr::rename(actor=id, actor_power=power, actor_domec=dom_ec, actor_rank=rank) %>%
  dplyr::rename(id=subject) %>%
  left_join(ec_rank, by=c("id", "bin")) %>%
  dplyr::rename(subject=id, subject_power=power, subject_domec=dom_ec, subject_rank=rank) %>%
  mutate(rule = if_else(actor_rank < subject_rank, "follower", "breaker"),# low dom_ec/rank = higher ranking
         color= if_else(rule=="breaker", "#f03b20","#2c7fb8")) %>% #"#dc8580", "#83b2d0" -> too dull
  arrange(-actor_power) 
glimpse(attrib_edge)
```

## visualize aggression networks

### network graph (20 birds)


First, for all 20 birds, where focals are highlighted
bin20birds <- c(1,4,5,8,9,12)

```{r aggression network 20 birds (highlight focal)}
  # choose bin
  run.code<-sort(unique(attrib_edge$bin))
  run.code
  
  # create graph from dataframe
 gdf <- graph.data.frame(attrib_edge)
 
  # create layout
 ## add rank 
 power_plot <- ec_rank %>% 
  filter(bin==run.code) %>%
  left_join(attrib_node) %>%
  dplyr::select(id, sex, power, rank) %>%
  arrange(-power) # changed to descending order rank instead of alphabethical order
 
 range(ec_rank$power)
 #range_ec_GPG <- range(ec_rank$dom_ec)

  treelayout <- layout_as_tree(gdf)  #V1=x, V2=y
  treelayout

  # change per bin
  ## change y axis to represent highest ranking on top of the netwwork
  ec_lay <- cbind(treelayout, power_plot$power) # observed ec rank 
  ec_lay<-ec_lay[,-2] # delete column 2

  ## change x axis to provide jitter of nodes
  #jitter<- c(0,-4,-5,1,2,3,0.5,3.5,-1.5,-3,-4,5,1.5,-2,2.5,5,-3.5,-1,-2.5,4.5,4)
  jitter20<- c(0,2,0.5,-3,-2,4,2.5,-4,3.5,-1,4.5,3,-4,-5,-0.5,-3,5,1,-2.5, 0) # diamond shape
  ec_lay <- cbind(jitter20, ec_lay)
  ec_lay<-ec_lay[,-2] # delete column 2
  #colnames(ec_lay) <- c("V1", "V2")

  
## set edge color
E(gdf)$rule <- as.character(attrib_edge$rule)
# unique(bin.data$color)
#E(gdf)$rule <- adjustcolor(as.character(bin.data$color),alpha=.6)

## add vertex/node attribute
V(gdf)$name <- power_plot$id
V(gdf)$dom <- power_plot$power # now we have to reverse y-axis
V(gdf)$sex <- power_plot$sex

## set node color  
# all green during removal period focals
#V(gdf)$fill <- c(alpha("#b8d79c", 1)) # all birds green

# highlight focal; "#1f78b4" for BBB; "#33a02c" for GPG; "#ff7f00" for OPP
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#1f78b4")[1+(V(gdf)$name=="BBB")]
V(gdf)$fill<- c(alpha("#b8d79c", 1), "#6a51a3")[1+(V(gdf)$name=="GPG")]  # "#6a51a3" -> PURPLE & "#33a02c" -> DARK GREEN
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#ff7f00")[1+(V(gdf)$name=="OPP")] # "#ff7f00" -> orange & 


 # plot graph
 #library(ggraph); class(gdf); gdf

set.seed(42)

# layout
layout <- create_layout(gdf, layout=ec_lay)
head(layout)


# plot

p8 <-
  ggraph(gdf, layout = ec_lay) +
  
  # --- edges
  # geom_edge_link0(aes(edge_width = weight), edge_colour = E(gdf)$color) + # edges going in multiple directions are overlapped
  geom_edge_fan(aes(edge_width = weight, 
                    colour=rule,
                    alpha = 0.6 # ..index..
                    ), 
                strength = 0.05) + # 0.05; curved edges. strength > 1 will create wider fans while the reverse will make them more narrow. If there are no parallel edges it behaves like geom_edge_link() and draws a straight line, but if parallel edges exists it will spread them out as arcs with different curvature.
  #scale_edge_alpha('Edge direction', guide = 'edge_direction') + # have to include ..index..instead of alpha 0.6 to create gradient
  scale_y_continuous(limits = c(0.82, 1.0), breaks = c(0.85,0.90, 0.95, 1.00) ) +
  scale_edge_width_continuous(range = c(0.2,3)) + 
  scale_edge_colour_manual(values = c("#f03b20", "#2c7fb8")) +
  
  # --- nodes
  geom_node_point(aes( shape= sex,
                       fill = as.factor(layout$fill),
                       color = "white", size=10)) + 
  # highlight focal; "#1f78b4" for BBB; green("#33a02c") or purple (alpha("#7030A0", 0.6) for GPG; "#ff7f00" for OPP
  scale_fill_manual(name = "fill",
                     values=  c("#6a51a3", alpha("#b8d79c", 1))) + #c(  "#1f78b4", alpha("#b8d79c", 1)) /  "#6a51a3" GPG / alpha("#b8d79c", 1), "#ff7f00") OPP
  scale_color_manual(values="white") +
  scale_shape_manual( name = "sex", 
                      values = c( 21, 23 ), #c( 21, 23 )  19, 17 
                      labels = c( "female", "male" )) +
  #geom_node_text(aes(label = layout$name), size=3) + #, , size=2 # bird names
  
  # --- layout
  #scale_size_continuous(range = c(1,6))+
  theme_graph()+ 
  labs(y="Power score" ) +
  ggtitle(paste("bin",run.code, sep=" ")) +
  theme(legend.position = "none", #"bottom"
        text=element_text(size=12),
        axis.line.y = element_line(),
        axis.text.y = element_text(),
        axis.title.y = element_text(),
        axis.ticks.y = element_line())  

 p8

 #ggsave(paste0("./figures/aggnwXbin_bin",run.code,"_3x3.pdf"), 
#        width = 3, height = 3,
#       device = cairo_pdf,
#       plot=p) #device = cairo_pdf,

# ggsave(paste0("./figures/aggnwXbin_bin",run.code,".pdf"), 
#       device = cairo_pdf,
#       plot=p) 
#dev.off()


#}

 
 nw_before_removal <- ggarrange(p1, p5, p9,  
          nrow = 3, ncol = 1)
  nw_before_removal
 ggsave( "./figures/nw_before_removal.pdf",
        width = 3.5, height = 9,
        device = cairo_pdf)


nw_after_reintro <- ggarrange(p4, p8, p12, 
                               nrow = 3, ncol = 1)
nw_after_reintro
ggsave( "./figures/nw_after==reintro.pdf",
        width = 3.5, height = 9,
        device = cairo_pdf)
```

### network graph (19 birds)
Second, for 19 birds, where focals are removed; this affects the layout. 
bin19birds <- c(2,3,6,7,10,11) 

```{r aggression network 19 birds (removal period)}

# combine dom_ec and rank to edge  dataframe
attrib_edge <- dyad.totalaggXbin %>%
  dplyr::select(-X) %>%
  #filter(bin %in% bintrial2_nofocal) %>%   # change 
  filter(bin %in% c(2,3,6,7,10,11) ) %>%
  #slice(rep(seq_len(n()), n.aggXbin)) %>%  # transform to edge list
  #mutate(weight = 1) %>%
  dplyr::rename(id=actor, weight=n.aggXbin) %>% 
  left_join(ec_rank, by=c("id", "bin")) %>% 
  dplyr::rename(actor=id, actor_power=power, actor_domec=dom_ec, actor_rank=rank) %>%
  dplyr::rename(id=subject) %>%
  left_join(ec_rank, by=c("id", "bin")) %>%
  dplyr::rename(subject=id, subject_power=power, subject_domec=dom_ec, subject_rank=rank) %>%
  mutate(rule = if_else(actor_rank < subject_rank, "follower", "breaker"),# low dom_ec/rank = higher ranking
         color= if_else(rule=="breaker", "#f03b20","#2c7fb8")) %>% #"#dc8580", "#83b2d0" -> too dull
  arrange(-actor_power) 
glimpse(attrib_edge)

  
loopby<-sort(unique(attrib_edge$bin))

for(i in seq_along(sort(unique(attrib_edge$bin)))){
  run.code <- 11
  print(run.code)
  
# SUBSET data by bin
 bin.data <- attrib_edge %>% 
   filter( bin==run.code) %>% #run.code
   arrange(-actor_power) 

  ## create graph from dataframe
 gdf <- graph.data.frame(bin.data)
 
  # create layout
 ## add rank 
 power_plot <- ec_rank %>% 
  filter(bin==run.code) %>%
  left_join(attrib_node) %>%
  dplyr::select(id, sex, power, rank) %>%
  arrange(-power) # changed to descending order rank instead of alphabethical order
 
 range(ec_rank$power)
 #range_ec_GPG <- range(ec_rank$dom_ec)

 
  treelayout <- layout_as_tree(gdf)  #V1=x, V2=y
  treelayout

  # change per bin
  ## change y axis to represent highest ranking on top of the netwwork
  ec_lay <- cbind(treelayout, power_plot$power) # observed ec rank 
  ec_lay<-ec_lay[,-2] # delete column 2

  ## change x axis to provide jitter of nodes
  #jitter<- c(0,-4,-5,1,2,3,0.5,3.5,-1.5,-3,-4,5,1.5,-2,2.5,5,-3.5,-1,-2.5,4.5,4)
  #jitter1<- c(0,2,0.5,3,5,4,2,3,4,-4,-1,-3,-4,-5,0,-3,-2,1,-2) # diamond shape
  jitter19<- c(0,2,0.5,-3,-2,4,2.5,-4,3.5,-1,4.5,3,-4,-5,0,-3,5,1,-2.5)
  ec_lay <- cbind(jitter19, ec_lay)
  ec_lay<-ec_lay[,-2] # delete column 2
  #colnames(ec_lay) <- c("V1", "V2")

  
## set edge color
E(gdf)$rule <- as.character(bin.data$rule)
# unique(bin.data$color)
#E(gdf)$rule <- adjustcolor(as.character(bin.data$color),alpha=.6)

## add vertex/node attribute
V(gdf)$name <- power_plot$id
V(gdf)$dom <- power_plot$power # now we have to reverse y-axis
V(gdf)$sex <- power_plot$sex

## set node color  
# all green during removal period focals
V(gdf)$fill <- c(alpha("#b8d79c", 1)) # all birds green

# highlight focal; "#1f78b4" for BBB; "#33a02c" for GPG; "#ff7f00" for OPP
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#1f78b4")[1+(V(gdf)$name=="BBB")]
#V(gdf)$fill<- c(alpha("#b8d79c", 1), "#6a51a3")[1+(V(gdf)$name=="GPG")]  # "#6a51a3" -> PURPLE & "#33a02c" -> DARK GREEN
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#ff7f00")[1+(V(gdf)$name=="OPP")] # "#ff7f00" -> orange & 


 # plot graph
 #library(ggraph); class(gdf); gdf

set.seed(42)

# layout
layout <- create_layout(gdf, layout=ec_lay)
head(layout)


# plot

p11 <-
  ggraph(gdf, layout = ec_lay) +
  
  # --- edges
  # geom_edge_link0(aes(edge_width = weight), edge_colour = E(gdf)$color) + # edges going in multiple directions are overlapped
  geom_edge_fan(aes(edge_width = weight, 
                    colour=rule,
                    alpha = 0.6 # ..index..
                    ), 
                strength = 0.05) + # 0.05; curved edges. strength > 1 will create wider fans while the reverse will make them more narrow. If there are no parallel edges it behaves like geom_edge_link() and draws a straight line, but if parallel edges exists it will spread them out as arcs with different curvature.
  #scale_edge_alpha('Edge direction', guide = 'edge_direction') + # have to include ..index..instead of alpha 0.6 to create gradient
  scale_y_continuous(limits = c(0.82, 1.0), breaks = c(0.85,0.90, 0.95, 1.00) ) +
  scale_edge_width_continuous(range = c(0.2,3)) + 
  scale_edge_colour_manual(values = c("#f03b20", "#2c7fb8")) +
  
  # --- nodes
  geom_node_point(aes( shape= sex,
                       fill = as.factor(layout$fill),
                       color = "white", size=10)) + 
  # highlight focal; "#1f78b4" for BBB; green("#33a02c") or purple (alpha("#7030A0", 0.6) for GPG; "#ff7f00" for OPP
  scale_fill_manual(name = "fill",
                     values=  c( alpha("#b8d79c", 1))) + #c(  "#1f78b4", alpha("#b8d79c", 1)) / alpha("#b8d79c", 1)) /  "#6a51a3" GPG
  scale_color_manual(values="white") +
  scale_shape_manual( name = "sex", 
                      values = c( 21, 23 ), #c( 21, 23 )  19, 17 
                      labels = c( "female", "male" )) +
  #geom_node_text(aes(label = layout$name), size=3) + #, , size=2 # bird names
  
  # --- layout
  #scale_size_continuous(range = c(1,6))+
  theme_graph()+ 
  labs(y="Power score" ) +
  ggtitle(paste("bin",run.code, sep=" ")) +
  theme(legend.position = "none", #"bottom"
        text=element_text(size=12),
        axis.line.y = element_line(),
        axis.text.y = element_text(),
        axis.title.y = element_text(),
        axis.ticks.y = element_line())  

 p11

 ggsave(paste0("./figures/aggnwXbin_bin",run.code,"_3.5x3.pdf"), 
        width = 3.5, height = 3,
       device = cairo_pdf,
       plot=p) #device = cairo_pdf,

 ggsave(paste0("./figures/aggnwXbin_bin",run.code,".pdf"), 
       device = cairo_pdf,
       plot=p) 
#dev.off()


}



```



## Network summary

```{r network summaries}
# make empty data frame to put data in
summ_network <- data.frame(run.code=character(),
                      density=numeric(),
                      total_agg =numeric(),
                      rulebreakers=numeric(),
                      rulefollowers=numeric())

  loopby<-sort(unique(attrib_edge$bin))
#i=1
  run.code <- loopby[i]
  print(run.code)
  
 # total aggressive events
 total_agg <- sum(attrib_edge$weight)

 # percent rule breakers and followers of total dyads
 rulebreakers <- length(attrib_edge$rule[attrib_edge$rule=="breaker"])/length(attrib_edge$actor) *100
 rulefollowers <- length(attrib_edge$rule[attrib_edge$rule=="follower"])/length(attrib_edge$actor) *100
 
 ## create graph from dataframe
 gdf <- graph.data.frame(attrib_edge)

  density <-   edge_density(gdf)

  # combine in dataframe
  pool <- cbind.data.frame(run.code, 
                         density,
                         total_agg,
                         rulebreakers,
                         rulefollowers)

  summ_network <-rbind(summ_network, pool )
 
  #summ_network1 <- summ_network
  #summ_network2 <- summ_network
  #summ_network3 <- summ_network
  #summ_network4 <- summ_network
  #summ_network5 <- summ_network
  #summ_network6 <- summ_network
  #summ_network7 <- summ_network
  #summ_network8 <- summ_network
  #summ_network9 <- summ_network
  #summ_network10 <- summ_network
  #summ_network11 <- summ_network
  #summ_network12 <- summ_network

# combine network summaries
summ_network <- bind_rows(summ_network1,
                          summ_network2,
                          summ_network3,
                          summ_network4,
                          summ_network5,
                          summ_network6,
                          summ_network7,
                          summ_network8,
                          summ_network9,
                          summ_network10,
                          summ_network11,
                          summ_network12)
summ_network <- summ_network %>%
  drop_na()

write.csv(summ_network, "output/network_summaries.csv")

```

```{r}
summ_network <- read.csv("output/network_summaries.csv") %>%
  dplyr::select(-X) %>%
  rename(period=run.code)

glimpse(summ_network)
write.csv(summ_network, "output/network_summaries_table.csv")

#summarize rules
mean(summ_network$rulefollowers)
sd(summ_network$rulefollowers)

mean(summ_network$rulebreakers)
sd(summ_network$rulebreakers)
```




## Mantel correlations
for aggression matrices:
  - before vs after removal
  - before vs after reintroduction

before vs after removal
BBB: bin c(1,2)
GPG: bin c(5,6)
OPP: bin c(9,10)

before reintroduction vs upon reintroduction 
BBB: bin  c(3,4) 
GPG: bin  c(7,8)
OPP: bin  c(11,12)

```{r correlation}

# CHANGE EACH TIME
observedXdyad <- dyad.totalaggXbin %>%
  filter(bin %in% c(11,12)) %>%
  pivot_wider(c(actor, subject), 
              names_from = bin, names_prefix = "bin",
              values_from = n.aggXbin,
              values_fill = 0) %>%
  dplyr::rename(before=bin11, after=bin12) # change each time
  
head(observedXdyad)
str(observedXdyad)

# check 
unique(observedXdyad$actor)
unique(observedXdyad$subject)


## correlation matrix  
# merge data with full dyad list in case any didn't interact
observedXdyad.alldyads <- merge(observedXdyad, dyad.list, 
                           by=c("actor", "subject"), 
                           all.y=TRUE)
head(observedXdyad.alldyads)
str(observedXdyad.alldyads) 

# convert NA to 0's
observedXdyad.alldyads[is.na(observedXdyad.alldyads)] <- 0

#check
head(observedXdyad.alldyads)
length(observedXdyad.alldyads$dyadID) # with 20 birds, we should have 380 dyads total

# before matrix
  obs.behavior1.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="before") #head(ref.behavior1.mx)
  obs.behavior1.mx[is.na(obs.behavior1.mx)] <- 0
  obs.behavior1.mx <- matrix.please(obs.behavior1.mx)
  
  # after matrix
  obs.behavior2.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="after") #head(ref.behavior2.mx)
  obs.behavior2.mx[is.na(obs.behavior2.mx)] <- 0
  obs.behavior2.mx <- matrix.please(obs.behavior2.mx)
  
  # Mantel test
  #mantel.test(obs.behavior1.mx, obs.behavior2.mx) # ape package
  obs.cor.result <- mantel(obs.behavior1.mx, obs.behavior2.mx, method = "spearman") # vegan package
  obs.cor.matrix <- obs.cor.result$statistic
obs.cor.result$statistic
obs.cor.result$signif
```

before vs after removal
BBB: bin c(1,2) -> mantel = 0.69, p = 0.001
GPG: bin c(5,6) -> mantel = 0.65, p = 0.001
OPP: bin c(9,10) -> mantel = 0.66, p = 0.001

before reintroduction vs upon reintroduction 
BBB: bin  c(3,4) -> mantel = 0.66, p = 0.001
GPG: bin  c(7,8) -> mantel = 0.51, p = 0.001
OPP: bin  c(11,12) -> mantel = 0.71, p = 0.001



# 2022 Aggression networks

join dominance scores to dyad aggression dataframe for both actor and subject

Structure of hierarchy:
- % of fights rule breakers vs rule followers

plot aggression network by dominance rank on y-axis and jitter on x-axis. One convenient way to adjust node coordinates is to plot with tkplot first, handtune the placement of the vertices, query the coordinates by the tk_coords function and use them with plot to plot the graph to any R device.

## bin selection

```{r}

bin_10birds <- c(3,4,7,8,11,12,15,16, 19,20,23,24)  # 
bin_11birds <- c(2,5,6, 9,10, 13,14, 17,18, 21,22, 25)
bin_reintro22_top <- c(5, 17, 25)
bin_reintro22_low <- c(9, 13, 21)

# per trial and period separate -> differently colored focals compared to the rest of the birds
bintrial1_focal <- c(2,5)
bintrial1_nofocal <- c(3,4)
bintrial2_focal <- c(6,9)
bintrial2_nofocal <- c(7,8)
bintrial3_focal <- c(10,13)
bintrial3_nofocal <- c(11,12)
bintrial4_focal <- c(14,17)
bintrial4_nofocal <- c(15,16)
bintrial5_focal <- c(18,21)
bintrial5_nofocal <- c(19,20)
bintrial6_focal <- c(22,25)
bintrial6_nofocal <- c(23,24)

```

## per social group

```{r}

```


## create edge list
change bin each time as the loop didn't work
```{r}
# check data
glimpse(aggXbin22)

aggXbin22$bin <- as.factor(aggXbin22$bin)

# combine dom_ec and rank to edge  dataframe
attrib_edge <- aggXbin22 %>%
  #filter(bin %in% bintrial2_nofocal) %>%   # change 
  filter(bin==2) %>%
  #slice(rep(seq_len(n()), n.aggXbin)) %>%  # transform to edge list
  #mutate(weight = 1) %>%
  dplyr::rename(id=actor, weight=n.aggXbin) %>% 
  left_join(power22, by=c("id", "bin")) %>% 
  dplyr::rename(actor=id, actor_power=power, actor_domec=dom_ec, actor_rank=rank) %>%
  dplyr::rename(id=subject) %>%
  left_join(power22, by=c("id", "bin")) %>%
  dplyr::rename(subject=id, subject_power=power, subject_domec=dom_ec, subject_rank=rank) %>%
  mutate(rule = if_else(actor_rank < subject_rank, "follower", "breaker"),# low dom_ec/rank = higher ranking
         color= if_else(rule=="breaker", "#f03b20","#2c7fb8")) %>% #"#dc8580", "#83b2d0" -> too dull
  arrange(-actor_power) 
glimpse(attrib_edge)
```

## visualize aggression networks

### network graph (20 birds)


First, for all 20 birds, where focals are highlighted
bin20birds <- c(1,4,5,8,9,12)

```{r aggression network 20 birds (highlight focal)}
  # choose bin
  run.code<-sort(unique(attrib_edge$bin))
  run.code
  
  # create graph from dataframe
 gdf <- graph.data.frame(attrib_edge)
 
  # create layout
 ## add rank 
 power_plot <- ec_rank %>% 
  filter(bin==run.code) %>%
  left_join(attrib_node) %>%
  dplyr::select(id, sex, power, rank) %>%
  arrange(-power) # changed to descending order rank instead of alphabethical order
 
 range(ec_rank$power)
 #range_ec_GPG <- range(ec_rank$dom_ec)

  treelayout <- layout_as_tree(gdf)  #V1=x, V2=y
  treelayout

  # change per bin
  ## change y axis to represent highest ranking on top of the netwwork
  ec_lay <- cbind(treelayout, power_plot$power) # observed ec rank 
  ec_lay<-ec_lay[,-2] # delete column 2

  ## change x axis to provide jitter of nodes
  #jitter<- c(0,-4,-5,1,2,3,0.5,3.5,-1.5,-3,-4,5,1.5,-2,2.5,5,-3.5,-1,-2.5,4.5,4)
  jitter20<- c(0,2,0.5,-3,-2,4,2.5,-4,3.5,-1,4.5,3,-4,-5,-0.5,-3,5,1,-2.5, 0) # diamond shape
  ec_lay <- cbind(jitter20, ec_lay)
  ec_lay<-ec_lay[,-2] # delete column 2
  #colnames(ec_lay) <- c("V1", "V2")

  
## set edge color
E(gdf)$rule <- as.character(attrib_edge$rule)
# unique(bin.data$color)
#E(gdf)$rule <- adjustcolor(as.character(bin.data$color),alpha=.6)

## add vertex/node attribute
V(gdf)$name <- power_plot$id
V(gdf)$dom <- power_plot$power # now we have to reverse y-axis
V(gdf)$sex <- power_plot$sex

## set node color  
# all green during removal period focals
#V(gdf)$fill <- c(alpha("#b8d79c", 1)) # all birds green

# highlight focal; "#1f78b4" for BBB; "#33a02c" for GPG; "#ff7f00" for OPP
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#1f78b4")[1+(V(gdf)$name=="BBB")]
V(gdf)$fill<- c(alpha("#b8d79c", 1), "#6a51a3")[1+(V(gdf)$name=="GPG")]  # "#6a51a3" -> PURPLE & "#33a02c" -> DARK GREEN
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#ff7f00")[1+(V(gdf)$name=="OPP")] # "#ff7f00" -> orange & 


 # plot graph
 #library(ggraph); class(gdf); gdf

set.seed(42)

# layout
layout <- create_layout(gdf, layout=ec_lay)
head(layout)


# plot

p8 <-
  ggraph(gdf, layout = ec_lay) +
  
  # --- edges
  # geom_edge_link0(aes(edge_width = weight), edge_colour = E(gdf)$color) + # edges going in multiple directions are overlapped
  geom_edge_fan(aes(edge_width = weight, 
                    colour=rule,
                    alpha = 0.6 # ..index..
                    ), 
                strength = 0.05) + # 0.05; curved edges. strength > 1 will create wider fans while the reverse will make them more narrow. If there are no parallel edges it behaves like geom_edge_link() and draws a straight line, but if parallel edges exists it will spread them out as arcs with different curvature.
  #scale_edge_alpha('Edge direction', guide = 'edge_direction') + # have to include ..index..instead of alpha 0.6 to create gradient
  scale_y_continuous(limits = c(0.82, 1.0), breaks = c(0.85,0.90, 0.95, 1.00) ) +
  scale_edge_width_continuous(range = c(0.2,3)) + 
  scale_edge_colour_manual(values = c("#f03b20", "#2c7fb8")) +
  
  # --- nodes
  geom_node_point(aes( shape= sex,
                       fill = as.factor(layout$fill),
                       color = "white", size=10)) + 
  # highlight focal; "#1f78b4" for BBB; green("#33a02c") or purple (alpha("#7030A0", 0.6) for GPG; "#ff7f00" for OPP
  scale_fill_manual(name = "fill",
                     values=  c("#6a51a3", alpha("#b8d79c", 1))) + #c(  "#1f78b4", alpha("#b8d79c", 1)) /  "#6a51a3" GPG / alpha("#b8d79c", 1), "#ff7f00") OPP
  scale_color_manual(values="white") +
  scale_shape_manual( name = "sex", 
                      values = c( 21, 23 ), #c( 21, 23 )  19, 17 
                      labels = c( "female", "male" )) +
  #geom_node_text(aes(label = layout$name), size=3) + #, , size=2 # bird names
  
  # --- layout
  #scale_size_continuous(range = c(1,6))+
  theme_graph()+ 
  labs(y="Power score" ) +
  ggtitle(paste("bin",run.code, sep=" ")) +
  theme(legend.position = "none", #"bottom"
        text=element_text(size=12),
        axis.line.y = element_line(),
        axis.text.y = element_text(),
        axis.title.y = element_text(),
        axis.ticks.y = element_line())  

 p8

 #ggsave(paste0("./figures/aggnwXbin_bin",run.code,"_3x3.pdf"), 
#        width = 3, height = 3,
#       device = cairo_pdf,
#       plot=p) #device = cairo_pdf,

# ggsave(paste0("./figures/aggnwXbin_bin",run.code,".pdf"), 
#       device = cairo_pdf,
#       plot=p) 
#dev.off()


#}

 
 nw_before_removal <- ggarrange(p1, p5, p9,  
          nrow = 3, ncol = 1)
  nw_before_removal
 ggsave( "./figures/nw_before_removal.pdf",
        width = 3.5, height = 9,
        device = cairo_pdf)


nw_after_reintro <- ggarrange(p4, p8, p12, 
                               nrow = 3, ncol = 1)
nw_after_reintro
ggsave( "./figures/nw_after==reintro.pdf",
        width = 3.5, height = 9,
        device = cairo_pdf)
```

### network graph (19 birds)
Second, for 19 birds, where focals are removed; this affects the layout. 
bin19birds <- c(2,3,6,7,10,11) 

```{r aggression network 19 birds (removal period)}

# combine dom_ec and rank to edge  dataframe
attrib_edge <- dyad.totalaggXbin %>%
  dplyr::select(-X) %>%
  #filter(bin %in% bintrial2_nofocal) %>%   # change 
  filter(bin %in% c(2,3,6,7,10,11) ) %>%
  #slice(rep(seq_len(n()), n.aggXbin)) %>%  # transform to edge list
  #mutate(weight = 1) %>%
  dplyr::rename(id=actor, weight=n.aggXbin) %>% 
  left_join(ec_rank, by=c("id", "bin")) %>% 
  dplyr::rename(actor=id, actor_power=power, actor_domec=dom_ec, actor_rank=rank) %>%
  dplyr::rename(id=subject) %>%
  left_join(ec_rank, by=c("id", "bin")) %>%
  dplyr::rename(subject=id, subject_power=power, subject_domec=dom_ec, subject_rank=rank) %>%
  mutate(rule = if_else(actor_rank < subject_rank, "follower", "breaker"),# low dom_ec/rank = higher ranking
         color= if_else(rule=="breaker", "#f03b20","#2c7fb8")) %>% #"#dc8580", "#83b2d0" -> too dull
  arrange(-actor_power) 
glimpse(attrib_edge)

  
loopby<-sort(unique(attrib_edge$bin))

for(i in seq_along(sort(unique(attrib_edge$bin)))){
  run.code <- 11
  print(run.code)
  
# SUBSET data by bin
 bin.data <- attrib_edge %>% 
   filter( bin==run.code) %>% #run.code
   arrange(-actor_power) 

  ## create graph from dataframe
 gdf <- graph.data.frame(bin.data)
 
  # create layout
 ## add rank 
 power_plot <- ec_rank %>% 
  filter(bin==run.code) %>%
  left_join(attrib_node) %>%
  dplyr::select(id, sex, power, rank) %>%
  arrange(-power) # changed to descending order rank instead of alphabethical order
 
 range(ec_rank$power)
 #range_ec_GPG <- range(ec_rank$dom_ec)

 
  treelayout <- layout_as_tree(gdf)  #V1=x, V2=y
  treelayout

  # change per bin
  ## change y axis to represent highest ranking on top of the netwwork
  ec_lay <- cbind(treelayout, power_plot$power) # observed ec rank 
  ec_lay<-ec_lay[,-2] # delete column 2

  ## change x axis to provide jitter of nodes
  #jitter<- c(0,-4,-5,1,2,3,0.5,3.5,-1.5,-3,-4,5,1.5,-2,2.5,5,-3.5,-1,-2.5,4.5,4)
  #jitter1<- c(0,2,0.5,3,5,4,2,3,4,-4,-1,-3,-4,-5,0,-3,-2,1,-2) # diamond shape
  jitter19<- c(0,2,0.5,-3,-2,4,2.5,-4,3.5,-1,4.5,3,-4,-5,0,-3,5,1,-2.5)
  ec_lay <- cbind(jitter19, ec_lay)
  ec_lay<-ec_lay[,-2] # delete column 2
  #colnames(ec_lay) <- c("V1", "V2")

  
## set edge color
E(gdf)$rule <- as.character(bin.data$rule)
# unique(bin.data$color)
#E(gdf)$rule <- adjustcolor(as.character(bin.data$color),alpha=.6)

## add vertex/node attribute
V(gdf)$name <- power_plot$id
V(gdf)$dom <- power_plot$power # now we have to reverse y-axis
V(gdf)$sex <- power_plot$sex

## set node color  
# all green during removal period focals
V(gdf)$fill <- c(alpha("#b8d79c", 1)) # all birds green

# highlight focal; "#1f78b4" for BBB; "#33a02c" for GPG; "#ff7f00" for OPP
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#1f78b4")[1+(V(gdf)$name=="BBB")]
#V(gdf)$fill<- c(alpha("#b8d79c", 1), "#6a51a3")[1+(V(gdf)$name=="GPG")]  # "#6a51a3" -> PURPLE & "#33a02c" -> DARK GREEN
#V(gdf)$fill<- c( alpha("#b8d79c", 1), "#ff7f00")[1+(V(gdf)$name=="OPP")] # "#ff7f00" -> orange & 


 # plot graph
 #library(ggraph); class(gdf); gdf

set.seed(42)

# layout
layout <- create_layout(gdf, layout=ec_lay)
head(layout)


# plot

p11 <-
  ggraph(gdf, layout = ec_lay) +
  
  # --- edges
  # geom_edge_link0(aes(edge_width = weight), edge_colour = E(gdf)$color) + # edges going in multiple directions are overlapped
  geom_edge_fan(aes(edge_width = weight, 
                    colour=rule,
                    alpha = 0.6 # ..index..
                    ), 
                strength = 0.05) + # 0.05; curved edges. strength > 1 will create wider fans while the reverse will make them more narrow. If there are no parallel edges it behaves like geom_edge_link() and draws a straight line, but if parallel edges exists it will spread them out as arcs with different curvature.
  #scale_edge_alpha('Edge direction', guide = 'edge_direction') + # have to include ..index..instead of alpha 0.6 to create gradient
  scale_y_continuous(limits = c(0.82, 1.0), breaks = c(0.85,0.90, 0.95, 1.00) ) +
  scale_edge_width_continuous(range = c(0.2,3)) + 
  scale_edge_colour_manual(values = c("#f03b20", "#2c7fb8")) +
  
  # --- nodes
  geom_node_point(aes( shape= sex,
                       fill = as.factor(layout$fill),
                       color = "white", size=10)) + 
  # highlight focal; "#1f78b4" for BBB; green("#33a02c") or purple (alpha("#7030A0", 0.6) for GPG; "#ff7f00" for OPP
  scale_fill_manual(name = "fill",
                     values=  c( alpha("#b8d79c", 1))) + #c(  "#1f78b4", alpha("#b8d79c", 1)) / alpha("#b8d79c", 1)) /  "#6a51a3" GPG
  scale_color_manual(values="white") +
  scale_shape_manual( name = "sex", 
                      values = c( 21, 23 ), #c( 21, 23 )  19, 17 
                      labels = c( "female", "male" )) +
  #geom_node_text(aes(label = layout$name), size=3) + #, , size=2 # bird names
  
  # --- layout
  #scale_size_continuous(range = c(1,6))+
  theme_graph()+ 
  labs(y="Power score" ) +
  ggtitle(paste("bin",run.code, sep=" ")) +
  theme(legend.position = "none", #"bottom"
        text=element_text(size=12),
        axis.line.y = element_line(),
        axis.text.y = element_text(),
        axis.title.y = element_text(),
        axis.ticks.y = element_line())  

 p11

 ggsave(paste0("./figures/aggnwXbin_bin",run.code,"_3.5x3.pdf"), 
        width = 3.5, height = 3,
       device = cairo_pdf,
       plot=p) #device = cairo_pdf,

 ggsave(paste0("./figures/aggnwXbin_bin",run.code,".pdf"), 
       device = cairo_pdf,
       plot=p) 
#dev.off()


}



```



## Network summary

```{r network summaries}
# make empty data frame to put data in
summ_network <- data.frame(run.code=character(),
                      density=numeric(),
                      total_agg =numeric(),
                      rulebreakers=numeric(),
                      rulefollowers=numeric())

  loopby<-sort(unique(attrib_edge$bin))
#i=1
  run.code <- loopby[i]
  print(run.code)
  
 # total aggressive events
 total_agg <- sum(attrib_edge$weight)

 # percent rule breakers and followers of total dyads
 rulebreakers <- length(attrib_edge$rule[attrib_edge$rule=="breaker"])/length(attrib_edge$actor) *100
 rulefollowers <- length(attrib_edge$rule[attrib_edge$rule=="follower"])/length(attrib_edge$actor) *100
 
 ## create graph from dataframe
 gdf <- graph.data.frame(attrib_edge)

  density <-   edge_density(gdf)

  # combine in dataframe
  pool <- cbind.data.frame(run.code, 
                         density,
                         total_agg,
                         rulebreakers,
                         rulefollowers)

  summ_network <-rbind(summ_network, pool )
 
  #summ_network1 <- summ_network
  #summ_network2 <- summ_network
  #summ_network3 <- summ_network
  #summ_network4 <- summ_network
  #summ_network5 <- summ_network
  #summ_network6 <- summ_network
  #summ_network7 <- summ_network
  #summ_network8 <- summ_network
  #summ_network9 <- summ_network
  #summ_network10 <- summ_network
  #summ_network11 <- summ_network
  #summ_network12 <- summ_network

# combine network summaries
summ_network <- bind_rows(summ_network1,
                          summ_network2,
                          summ_network3,
                          summ_network4,
                          summ_network5,
                          summ_network6,
                          summ_network7,
                          summ_network8,
                          summ_network9,
                          summ_network10,
                          summ_network11,
                          summ_network12)
summ_network <- summ_network %>%
  drop_na()

write.csv(summ_network, "output/network_summaries.csv")

```

```{r}
summ_network <- read.csv("output/network_summaries.csv") %>%
  dplyr::select(-X) %>%
  rename(period=run.code)

glimpse(summ_network)
write.csv(summ_network, "output/network_summaries_table.csv")

#summarize rules
mean(summ_network$rulefollowers)
sd(summ_network$rulefollowers)

mean(summ_network$rulebreakers)
sd(summ_network$rulebreakers)
```




## Mantel correlations
for aggression matrices:
  - before vs after removal
  - before vs after reintroduction

before vs after removal
BBB: bin c(1,2)
GPG: bin c(5,6)
OPP: bin c(9,10)

before reintroduction vs upon reintroduction 
BBB: bin  c(3,4) 
GPG: bin  c(7,8)
OPP: bin  c(11,12)

```{r correlation}

# CHANGE EACH TIME
observedXdyad <- dyad.totalaggXbin %>%
  filter(bin %in% c(11,12)) %>%
  pivot_wider(c(actor, subject), 
              names_from = bin, names_prefix = "bin",
              values_from = n.aggXbin,
              values_fill = 0) %>%
  dplyr::rename(before=bin11, after=bin12) # change each time
  
head(observedXdyad)
str(observedXdyad)

# check 
unique(observedXdyad$actor)
unique(observedXdyad$subject)


## correlation matrix  
# merge data with full dyad list in case any didn't interact
observedXdyad.alldyads <- merge(observedXdyad, dyad.list, 
                           by=c("actor", "subject"), 
                           all.y=TRUE)
head(observedXdyad.alldyads)
str(observedXdyad.alldyads) 

# convert NA to 0's
observedXdyad.alldyads[is.na(observedXdyad.alldyads)] <- 0

#check
head(observedXdyad.alldyads)
length(observedXdyad.alldyads$dyadID) # with 20 birds, we should have 380 dyads total

# before matrix
  obs.behavior1.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="before") #head(ref.behavior1.mx)
  obs.behavior1.mx[is.na(obs.behavior1.mx)] <- 0
  obs.behavior1.mx <- matrix.please(obs.behavior1.mx)
  
  # after matrix
  obs.behavior2.mx <- reshape2::dcast(observedXdyad.alldyads, actor~subject, value.var="after") #head(ref.behavior2.mx)
  obs.behavior2.mx[is.na(obs.behavior2.mx)] <- 0
  obs.behavior2.mx <- matrix.please(obs.behavior2.mx)
  
  # Mantel test
  #mantel.test(obs.behavior1.mx, obs.behavior2.mx) # ape package
  obs.cor.result <- mantel(obs.behavior1.mx, obs.behavior2.mx, method = "spearman") # vegan package
  obs.cor.matrix <- obs.cor.result$statistic
obs.cor.result$statistic
obs.cor.result$signif
```

before vs after removal
BBB: bin c(1,2) -> mantel = 0.69, p = 0.001
GPG: bin c(5,6) -> mantel = 0.65, p = 0.001
OPP: bin c(9,10) -> mantel = 0.66, p = 0.001

before reintroduction vs upon reintroduction 
BBB: bin  c(3,4) -> mantel = 0.66, p = 0.001
GPG: bin  c(7,8) -> mantel = 0.51, p = 0.001
OPP: bin  c(11,12) -> mantel = 0.71, p = 0.001
