---
title: "Stats_dompatterns_start2021"
author: "Annemarie van der Marel"
date: '2022-09-23'
output: html_document
---

#Session info
```{r}
sessionInfo()

#citation('diagram')
citation('ggridges')
```


#Load packages
```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(reshape2)
library(stringr)
library(purrr)
library(domstruc) #install_github("danm0nster/domstruc")
library(ggraph)
library(diagram)
library(ggpubr)
library(ggridges)
library(igraph)
library(ggraph)
library(vegan)
```

# Import data
First, we have to run the "BehaveTable data import.Rmd" to compile the behavioral observations. 

I included the code to remove the duplicates here as we are only looking at displacements and crowds, but I also have that in a separate Rmd script ("cleanup interaction file.Rmd") to include the other interactions. 

Data files will be present in the Github repository, so we can link R to the repository
```{r data import, include=FALSE, echo=FALSE}
## raw data
behave_raw <- read.csv("./data/2021_interactions_raw.csv") %>%
  dplyr::select(-X)
unique(behave_raw$date)

behave <- behave_raw


sort(unique(behave$date))


# period file
#obs.day.key <- read.csv("../2021_timeline_fieldwork.csv") 
bin.dates <- read.csv("./data/2021_3daybins.csv") 
bin.dates$bin <- as.integer(bin.dates$bin)
bins <- bin.dates %>%
  #filter(bin<10) %>% 
  dplyr::select(date, type, bin, n_birds) 

bin_start <- bin.dates %>%
  filter(bin<11) %>% 
  dplyr::select(date, bin, type, n_birds) %>%
  group_by(bin) %>% 
  slice_tail() %>%
  ungroup()

bin_start_dates <- bin.dates %>%
  filter(bin<11) %>% 
  dplyr::select(date, bin, type, n_birds)


## bird list
birdIDs_capture<- read.csv("./data/2021_birdIDs.csv") 
# start
birdIDs <- birdIDs_capture %>%
  dplyr::select(mark_id) %>%
  filter(mark_id!="") # all 22 birds
birdIDs_removalstorm <- birdIDs %>% 
  filter(!mark_id %in% c("BBP", "OBG"))


# capture file
trap_raw<-read.csv("./data/2021_MOPA_capture file.csv")
glimpse(trap_raw)

trap <- trap_raw %>% 
  rename(id=mark_id)


# observation hours
obs_hours_raw <- read.csv("./data/2021_observation hours_cleaned.csv") %>%
  dplyr::select(-X)


```


# Ready data  
## change bird names for actor and subject
 (mark_id1 = mark_id2)

we then have to allocate a line for each of these birds with 0's to make the matrix complete
```{r}

## mark_id1 = mark_id2 ; from 22 to 20 birds
# GBB = NPB
behave$actor[behave$actor=="GBB"] <- "NPB"
behave$subject[behave$subject=="GBB"] <- "NPB"

# BOO = POO
behave$actor[behave$actor=="BOO"] <- "POO"
behave$subject[behave$subject=="BOO"] <- "POO"

## mark_id2 = mark_id3 ; different colors
# GBB = NPB
behave$actor[behave$actor=="NPB"] <- "PPB"
behave$subject[behave$subject=="NPB"] <- "PPB"


# check
unique(behave$actor)
unique(behave$subject)



```

##Include bin and period to behavior file + orderkey
2021-04-11 start of bin2 -> big thunderstorm during the night, many bird headwounds by flying into the chicken wire (upon capture on 2021-04-27, we had to remove 2  injured birds)


```{r}
behave_bin <- behave %>%
  left_join(bins)

unique(behave_bin$bin)
head(behave_bin)
tail(behave_bin)

# how many days per period
day_sum<- behave_bin %>% 
  group_by( bin) %>%
  summarise(n=n_distinct(date))

```

## node attribute list
```{r node attributes}
attrib_node <- birdIDs_capture %>%
  filter(mark_id!="") %>%
  mutate(id=mark_id) %>%
  dplyr::select(band_id, id, sex, site_captured, cage_id_2021)

```

## check for removed birds in social perturbation period
BBP
OBG

```{r}

unique(behave_bin$type)
period20birds <- c("duo", "switch")

check_social_actor<-filter(behave_bin, type %in% period20birds, 
                           actor == "BBP"|actor =="OBG")
check_social_subject<-filter(behave_bin, type %in% period20birds, 
                             subject == "BBP"|subject =="OBG")
check_social<-bind_rows(check_social_actor,check_social_subject)
length(behave_bin$orderkey)-length(check_social$orderkey) 

#behave_correct<-anti_join(behavekey, check_social, by = "orderkey")
#length(behave_correct$orderkey)
```
no instances of removed birds


## Helper functions & data
```{r echo=FALSE}
# include calibri as text font
windowsFonts(Calibri = windowsFont("Calibri"))
windowsFonts()


#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}


```


## bird list
change for each perturbation as different birds are removed

```{r bird list 22 birds}
birdIDs22 <- birdIDs_capture %>%
  dplyr::select(mark_id) %>%
  filter(mark_id!="") 

bird.list <- sort(unique(birdIDs22$mark_id)) # change each time
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

```


```{r bird list 20 birds}

## bin 7 onward -> 20 birds 

birdIDs20 <- birdIDs_capture %>%
  dplyr::select(mark_id3) %>%
  filter(mark_id3!="") # all 22 birds # change each time
bird.list <- sort(unique(birdIDs20$mark_id3))
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

```


# Remove duplicates
BY bin: for start of 2021 season -> bin 1-10; social manipulation experiment started at bin 10 which is rank assessment period 0

more conservative: removing duplicates using crowds and displacements combined from the beginning, results in less agonistic events but mistakes per behavior (entered crowd while it had to be displacement) are excluded. 

We started differentiating between aggression at the feeder and away from the feeder so we have to make certain to include both. 

```{r remove duplicates conservative method X bin}
#### Remove duplicates for displacements and crowds ####
##more conservative: removing duplicates by crowds and displacements combined
#behave_totalagg<-behave_totalagg_removal

# include all aggression (at the feeder and away from feeder)
unique(behave_bin$behavior)

# all aggression (both feeders and nonfeeders)
behave_agg <- behave_bin %>%
  separate(behavior, c("behavior", "type"), sep="_") # all aggression at both feeders and away from feeders

unique(behave_agg$behavior)
unique(behave_agg$date)

# filter displacement & crowd data from period prior to social perturbations ()
aggDC <- behave_agg %>% 
  filter(behavior=="displace" | behavior=="crowd", 
         bin %in% c(7, 8,9,10) # bin %in% c(7, 8,9,10) = 20 birds
         # bin<7 = 22 birds, 
         ) %>%
  dplyr::select(orderkey,  sessionKEY, type, bin, date, time, actor, subject, behavior)

str(aggDC)
length(aggDC$sessionKEY)
unique(aggDC$bin)

# keep only observations where the actor is in the list of bird IDs
aggDC.goodID <- subset(aggDC, actor %in% bird.list) # check size of bird list(20 or 22 birds)

# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
aggDC.goodID <- subset(aggDC.goodID, subject %in% bird.list)


## for total agonistic events (crowds and displacements combined)
#finds n CD and crowds by sessionKEY, date, time, actor, subject, behavior
dyad.totalagg.key <- aggDC.goodID %>% 
  dplyr::group_by(sessionKEY, bin, date, time,actor, subject) %>%  # removed , behavior
  tally() # counts behaviors that were observed within the same minute

#finds max agg per actor by date, time, and behavior type. Return just that summary
dyad.totalagg.maxkey <- dyad.totalagg.key %>% 
  group_by(actor, subject,bin,  date, time) %>% # removed behavior, behavior
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
#head(dyad.totalagg.maxkey)

#finds n disps per actor by date for summarized/trimmed data
dyad.totalaggXbin <- dyad.totalagg.maxkey %>% 
  group_by(actor, subject, bin) %>% # removed behavior to get total aggression
  summarise(n.aggXbin=sum(n)) %>% 
  ungroup()
#head(dyad.totalaggXbin)
sum(dyad.totalaggXbin$n.aggXbin)


#Here we are interested in total number of interactions instead of chronological #order of interactions. If you require order of interactions, then next steps will #be different from the following setup as you will have to add the behaviors with #n>1 as separate rows. 

# save summarized aggression X bin df
#write.csv(dyad.totalaggXbin, "./output/dyad.aggXbin.start21.csv")

```
combine datasets with 22 birds (bin<7) and 20 birds (bin 7:10)
```{r}
#dyad.totalaggXbin_22birds <- dyad.totalaggXbin
#dyad.totalaggXbin_20birds <- dyad.totalaggXbin

dyad.totalaggXbin <- bind_rows(dyad.totalaggXbin_22birds,
                               dyad.totalaggXbin_20birds)
#write.csv(dyad.totalaggXbin, "./output/dyad.aggXbin.start21.csv")

```


# Analyses

## Total aggression
```{r sum total aggression}
## Summarize displacements & crowds combined --> total aggression
totalagg<- read.csv("./output/dyad.aggXbin.start21.csv")
sum(totalagg$n.aggXbin) = sum(dyad.totalaggXbin_22birds$n.aggXbin) + sum(dyad.totalaggXbin_20birds$n.aggXbin)
# all good

str(dyad.totalaggXbin)

# check 
unique(dyad.totalaggXbin$actor)
length(unique(dyad.totalaggXbin$actor))
unique(dyad.totalaggXbin$subject)
length(unique(dyad.totalaggXbin$subject))
sort(unique(dyad.totalaggXbin$bin))

sum(dyad.totalaggXbin$n.aggXbin)

# summarize total aggression events per bin
sum.aggDCXbin_start21 <-dyad.totalaggXbin %>%
  group_by(bin) %>%
  summarize(n.agg = sum(n.aggXbin), 
            n.dyads=length(actor), 
            n.actors=n_distinct(actor),
            n.subjects=n_distinct(subject))
sum.aggDCXbin_start21

n.aggDC.start21 <- sum(sum.aggDCXbin_start21$n.agg)

# check 
sum(dyad.totalaggXbin$n.aggXbin)==sum(sum.aggDCXbin_start21$n.agg)




```
Before the social perturbation experiment in 2021, we observed 18858 interactions (excluding unknown individuals and duplicates)





## Social dominance pattern X 3-day bin

We have to run the dominance pattern and rank analysis separately for the removal periods as we excluded one of the group members. Therefore when we are making the matrix we have to filter out the top-ranked focal bird, which is a different ID each trial. 

```{r}

#check.df <- read.csv( "./output/dyad.aggXbin.start21.csv")
glimpse(dyad.totalaggXbin)

unique(dyad.totalaggXbin$actor)
unique(dyad.totalaggXbin$subject)
length(unique(dyad.totalaggXbin$actor))
length(unique(dyad.totalaggXbin$subject))


## for bins with all 22 birds ----
bin22birds<- c(1:6)
dyad.totalaggXbin22 <- dyad.totalaggXbin %>%
  filter(bin %in% bin22birds)
#dyad.totalaggXbin_22birds == dyad.totalaggXbin22

# change bird list
bird.list <- sort(unique(birdIDs22$mark_id)) # change for social period
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")


## social dominance pattern ----

dyad.totalaggXbin <- dyad.totalaggXbin22 %>%
    arrange( bin, by_group = FALSE) 


# make empty data frame to put data in
strategies <- data.frame(run.code=character(),
                      focus=numeric(),
                      position=numeric(),
                      strategy=character())

# run = 1

start.time <- Sys.time()

loopby<-unique(dyad.totalaggXbin$bin)

i=1

for(i in seq_along(unique(dyad.totalaggXbin$bin))){
  run.code <- loopby[i]
  print(run.code)
  run.data <- subset(dyad.totalaggXbin, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
  data_alldyads <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$date<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)

  # ---- aggression strategy  
  # Compute focus & position
  focus <- dom_focus(mx)
  position <- dom_position(mx)
  fp <- cbind.data.frame(focus, position)
  colnames(fp) <- c("focus", "position")
  
  #Compute blur models
  blur <- dom_make_blur_data(mx)
  dom_plot_strategy(fp, blur) # plot aggression stategies
  
  #Find strategy
  strategy <- dom_categorize_strategy(data=fp, blur_data=blur)
  
  # --- combine dataframe
pool <-  cbind.data.frame(run.code, 
                         focus, 
                         position, 
                         strategy) 

strategies <- rbind(pool, strategies)

}

end.time <- Sys.time()

#Time to run:
end.time - start.time

strategies_22birds <- strategies


```

This is the analyses for 20 birds
```{r Dominance pattern using crowds and displacements_ conservative}
## removal injured birds ----
# bin20birds <- c(7:10)
# dyad.totalaggXbinR1 <- dyad.totalaggXbin_20birds %>%
#   filter(bin %in% bin20birds)
# dyad.totalaggXbin_20birds == dyad.totalaggXbinR1
# unique(dyad.totalaggXbinR1$bin)

# change bird list
bird.list <- sort(unique(birdIDs_removalstorm$mark_id)) # BBP and OBG removed
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

unique(dyad.totalaggXbinR1$actor) # is BBP and OBG excluded?
unique(dyad.totalaggXbinR1$subject)
length(unique(dyad.totalaggXbinR1$actor))
length(unique(dyad.totalaggXbinR1$subject))

## social dominance pattern ----

dyad.totalaggXbin <- dyad.totalaggXbin_20birds %>%
    arrange( bin, by_group = FALSE) 


# make empty data frame to put data in
strategies <- data.frame(run.code=character(),
                      focus=numeric(),
                      position=numeric(),
                      strategy=character())

# run = 1

start.time <- Sys.time()

loopby<-unique(dyad.totalaggXbin$bin)

i=1

for(i in seq_along(unique(dyad.totalaggXbin$bin))){
  run.code <- loopby[i]
  print(run.code)
  run.data <- subset(dyad.totalaggXbin, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
  data_alldyads <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$date<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)

  # ---- aggression strategy  
  # Compute focus & position
  focus <- dom_focus(mx)
  position <- dom_position(mx)
  fp <- cbind.data.frame(focus, position)
  colnames(fp) <- c("focus", "position")
  
  #Compute blur models
  blur <- dom_make_blur_data(mx)
  dom_plot_strategy(fp, blur) # plot aggression stategies
  
  #Find strategy
  strategy <- dom_categorize_strategy(data=fp, blur_data=blur)
  
  # --- combine dataframe
pool <-  cbind.data.frame(run.code, 
                         focus, 
                         position, 
                         strategy) 

strategies <- rbind(pool, strategies)

}

end.time <- Sys.time()

#Time to run:
end.time - start.time

strategies_removalstorm <- strategies

```

```{r change each time}

# strategies_22birds <- strategies
# random_fights<- c(0.28, 0.22, 0.34, 0.29, 0.27, 0.26)
# strategies_22birds <- bind_cols(strategies_22birds, random_fights) %>%
#   rename(random_fights='...5')

# strategies_removalstorm <- strategies
# random_fights<- c(0.32, 0.39, 0.38)
# strategies_storm <- bind_cols(strategies_removalstorm, random_fights) %>%
#   rename(random_fights='...5')

strategies2021 <- bind_rows(strategies_22birds, 
                            strategies_removalstorm)
strategies2021 <- strategies2021 %>%
  rename(bin=run.code) %>%
  arrange(bin) %>%
  mutate(dominance_pattern= if_else(strategy=="downward.heuristic", "downward heuristic","bullying") ) %>%
  dplyr::select(-strategy)

write.csv(strategies2021, "dompatterns_start2021.csv")                                            
```



## individual ranks X bin
Which ranks were the 2 injured birds?


```{r indivdidual rank}

head(dyad.totalaggXbin22)
sort(unique(dyad.totalaggXbin22$bin))
sort(unique(dyad.totalaggXbinR1$bin))

dyad.aggXbin <- dyad.totalaggXbinR1

#### DOMINANCE RANK BY bin####

#head(dyad.totalaggXbin)

# make empty data frame to put data in
ec_rank <- data.frame(run.code=character(),
                      bird.list=character(), 
                      dom_ec=numeric(),
                      rank=numeric())

# run = 1

start.time <- Sys.time()

# select only bins prior removal
loopby<-unique(dyad.aggXbin$bin)


#i=1

for(i in seq_along(loopby)){
  run.code <- loopby[i]
  print(run.code)
  run.data <- subset(dyad.aggXbin, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
  data_alldyads <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 20 birds: 380 dyads, with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$date<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)
  
  ## dominance
  dom_ec <- dom_ec(mx)

  rank <- dom_ranks(dom_ec)


  # combine in dataframe
  pool <- cbind.data.frame(run.code, 
                         bird.list,
                         dom_ec,
                         rank)

  ec_rank<-rbind(ec_rank, pool )

}

end.time <- Sys.time()

#Time to run:
end.time - start.time

# check
glimpse(ec_rank)
length(unique(ec_rank$run.code))
check <- ec_rank %>% 
  group_by(run.code) %>% 
  summarize(length(bird.list))

ec_rankXbin <- ec_rank %>%
  rename(bin=run.code, id=bird.list) %>%
  mutate(power=1-dom_ec, 
         rank=21-rank) %>%
  group_by(bin) %>%
  arrange(bin, -power)




```


```{r}
# rank of removed injured birds

duo <- ec_rankXbin %>%
  filter(bin==6, id %in% c("BBP", "OBG"))
```

OBG was ranked 10
BBP was ranked 16

```{r}
# combine
#ec_rankXbin22 <- ec_rankXbin
ec_rankXbin20 <- ec_rankXbin

powerXbin <- bind_rows(ec_rankXbin22, ec_rankXbin20) %>%
  dplyr::select(-dom_ec)

write.csv(powerXbin, "./output/powerXbin_start.csv")

```


# Figures
## Fig 1c) aggression rate controlled for hours observed and group size

#### observation hours
Total hours is calculated without pee breaks as observations continued. Calculate for
  total hours per bin

```{r observation hours X bin}

glimpse(obs_hours_raw)

obs_hours_raw$breaks[obs_hours_raw$breaks=="break"] <- "lunch"

## lubridate: hm(..., quiet = FALSE, roll = FALSE)
# obs_hours$start <-  hm(obs_hours$start)
# obs_hours$end <-  hm(obs_hours$end)

## total observation hours
hoursobs <- obs_hours_raw %>%
  right_join(bin_start_dates, by="date") %>% # length(unique(hoursobs$date))
  filter(breaks %in% c("obs", "lunch")) %>%
  group_by(bin, breaks) %>%
  summarize(sum=sum(totaltime_hour)) %>%
  pivot_wider(names_from = breaks,
              values_from = sum) %>%
  mutate(hours_obs=obs-lunch) %>%
  dplyr::select(bin, hours_obs)


sum(hoursobs$hours_obs)

sumhoursobs <- hoursobs %>% 
  ungroup() %>%
  summarize(meanhoursobs=mean(hours_obs),
            sdhoursobs=sd(hours_obs),
            n=n(),
            se=sdhoursobs/sqrt(n))

```
Person hours  (remove breaks and multiply observation hours by the number of observers)


```{r person hours X bin}

## person hours

# total over 4 observed (lunch already excluded)
sumperson <- hoursobs %>%
  mutate(sumperson=hours_obs*4)

# personal breaks
person_hoursXbin <- obs_hours_raw %>%
  right_join(bin_start_dates, by="date") %>%
  filter(!subject %in% c("lunch","")) %>%
  group_by(bin) %>%
  summarise(breaks=sum(totaltime_hour, na.rm = T)) %>%
  left_join(sumperson) %>%
  mutate(person_hours = sumperson-breaks)

sum(person_hoursXbin$person_hours)

## combine
obs.hours <- hoursobs %>% 
  left_join(person_hoursXbin) %>%
  dplyr::select(bin, hours_obs, person_hours ) 


```

220.9 hours and 832.8 person hours across 31 days, with an average of 22.1 ± 2.4 (SD) hours of observation per three-day bin (n = 10 bins)


#### groups size and aggression rate X bin

```{r plot aggression controlled for group size and hours observed}


# total aggression
glimpse(sum.aggDCXbin_start21)

# mean total of agonistic events observed per bin
summ_agXbin <- sum.aggDCXbin_start21 %>%
  summarize(mean_ag =mean(n.agg),
            sd_ag = sd(n.agg))

# mean and SD value
bin_start_dates$n_birds <- as.integer(bin_start_dates$n_birds )
summ_ag_21start <- sum.aggDCXbin_start21 %>%
  left_join(bin_start_dates %>%
              group_by(bin) %>%
              slice_head()) %>%
  left_join(obs.hours, by="bin") %>%
  mutate(ag_bird= n.agg/n_birds, 
         ag_bird_obs= n.agg/n_birds/hours_obs, 
         ag_person_obs = n.agg/n_birds/person_hours, 
         mean = mean(ag_bird_obs),
         sd = sd(ag_bird_obs),
         n=n(),
         se=sd/sqrt(n))

ave.ag.rate <- mean(summ_ag_21start$ag_bird_obs)
sd.ag.rate <- sd(summ_ag_21start$ag_bird_obs)
min(summ_ag_21start$ag_bird_obs)
max(summ_ag_21start$ag_bird_obs)

# ave.ag.rate <- mean(summ_ag_21start$ag_person_obs)
# sd.ag.rate <- sd(summ_ag_21start$ag_person_obs)
# 
# min(summ_ag_21start$ag_person_obs)
# max(summ_ag_21start$ag_person_obs)
```

#### plot aggression rate

```{r}
plot_controlaggression_21start<-  summ_ag_21start %>%
  
  ggplot( aes(x = bin)) +
  
  # add average rate of aggression
  geom_hline(yintercept = ave.ag.rate, color = alpha("grey", 0.6), size=2) +
  
  # aggression controlled for group size and hours observed
  geom_line(aes(y=ag_bird_obs), size=1) + 
  geom_point(aes(y=ag_bird_obs), size=3) +
  
  scale_y_continuous(limits=c(0, 5.5),
                     breaks= c(1:5)) +
  scale_x_continuous(limits = c(1, 10), 
                     breaks= c(1:10)) +
  
  theme_classic() +
  theme(legend.position="bottom",
         text = element_text(size=14),
          axis.text = element_text(size=12),
         axis.title.x = element_blank()) +
   #     axis.text.x = element_blank(),
    #    axis.ticks.x=element_line()) +
  labs(y="aggression rate") + #, title = "Amount of aggression controlled for group size and hours observed"
  
 geom_vline(xintercept = 1.5, linetype="dashed", color =  alpha("black", 0.6)) + # thunderstorm
 geom_vline(xintercept = 6.5, linetype="dashed", color =  alpha("black", 0.6)) +# removal of 2 injured birds
  geom_vline(xintercept = 7.5, linetype="dashed", color =  alpha("black", 0.6)) # dominant partner switch
  
 
plot_controlaggression_21start
```

## Fig 1d)  dominance patterns
strategies take long to run, so import from file

```{r plot patterns over time}
strategies2021 <- read.csv("output/summ_agg2021.csv")
strategies <- strategies2021 %>%
  dplyr::select(bin, dominance_pattern) %>%
  #filter(bin!=4) %>%
  add_row( bin=21, dominance_pattern="bullying") %>%
  add_row( bin=25, dominance_pattern="close competitor") %>%
  mutate(rank = if_else(dominance_pattern=="downward heuristic", 1, 
                        ifelse(dominance_pattern=="bullying", 2, 3))) %>% 
  mutate(use = if_else(dominance_pattern=="downward heuristic", "A", 
                        ifelse(dominance_pattern=="bullying", "B", "C"))) %>% 
  mutate(observed_pattern = factor(dominance_pattern, 
            levels =c("bullying", "close competitor","downward heuristic"),
            labels =c("Bullying", "Close competitor","Downward heuristic"))) #unique(dominance_pattern))


# dominance patterns long perturbation period
plot_strategies_summ_start21 <- ggplot(strategies, 
                               aes(bin, observed_pattern)) +
  geom_point(size = 6, shape = 21, colour = "black", aes(fill = observed_pattern )) +
  labs(title = "", y="", x="social dominance pattern assessment periods") + #title = "social dominance pattern"
  scale_fill_manual(values = c( "#0000FF","#C00000" , "#7F7F7F")) +
  theme_classic() + #theme_minimal() 
  theme(#axis.title = element_text(size=12, angle = 45), #element_blank()
        #axis.text.y =element_text(size=12, angle = 25), 
        legend.position = "none",
        text = element_text(size =14),
        axis.text = element_text(size=12),
        plot.margin = unit(c(1, 1, 0.5, 1), "cm")) +
  
  scale_x_continuous(limits = c(1, 10),
                     breaks= c(1:10)) +
  
  # add lines to highlight perturbations

  geom_vline(xintercept = 1.5, linetype="dashed", color =  alpha("black", 0.6)) + # thunderstorm
  geom_vline(xintercept = 6.5, linetype="dashed", color =  alpha("black", 0.6)) + # removal 2 injured birds
geom_vline(xintercept = 7.5, linetype="dashed", color =  alpha("black", 0.6)) # dominant partner switch

plot_strategies_summ_start21

```

# Plot 2021 data
```{r}

  ggarrange(plot_controlaggression_21start, plot_strategies_summ_start21, 
            nrow=2, ncol=1, align = "hv",
            labels = c("(c)", "(d)")
            ) # legend = "bottom", label.x = 0.15, label.y=0.85

  ggsave("./figures/plot_Fig2021_start_hoursv1.pdf", width= 175, height=125, units="mm")
         #width= 7, height=7, units="in") # full page  width= 7, height=11, units="in") # half page width= 125, height=250, units="mm"
```


# write aggression summary


```{r}
agg_summ21 <- strategies2021 %>%
  left_join(summ_ag_all) %>%
   dplyr::select(-date)

agg_summ21$focus <- round(agg_summ21$focus , digits = 2)
agg_summ21$position <- round(agg_summ21$position , digits = 2)
agg_summ21$hours_obs <- round(agg_summ21$hours_obs , digits = 2)
agg_summ21$ag_bird <- round(agg_summ21$ag_bird , digits = 2)
agg_summ21$ag_bird_obs <- round(agg_summ21$ag_bird_obs , digits = 2)


write.csv(agg_summ21, "output/summ_agg2021.csv")

```




