---
title: "2020_dominance patterns"
author: "Annemarie van der Marel"
date: '2022-10-18'
output: html_document
---


R code to obtain the dominance hierarchy of 3-day period in 2020
  - social dominance pattern

Basics
-	Number of total aggressive events observed
-	Total hours observed

Figure
-	Daily amount of aggression received/given corrected for hours observed
-	Time series social dominance patterns
-	transition diagram observed and random expectation



Decisions to make for datasets:
- use crowds and displacements combined
- use aggression away from feeder or all aggression
- remove duplicates combined 

#Session info
```{r}
sessionInfo()

#citation('diagram')

```


#Load packages
```{r, include=FALSE, echo=FALSE}
library(tidyverse)
library(reshape2)
library(stringr)
library(purrr)
library(domstruc) #xfun::install_github("danm0nster/domstruc")
 library(ggraph)
library(diagram)
library(ggpubr)
library(ggridges)
library(igraph)
library(ggraph)
library(vegan)
library(lubridate)
```

# Import raw data


```{r}
behave_clean <- read.csv('./data/2020_interactions_all.csv') %>%
  dplyr::select(-X)
  
obs.hours <- read.csv('./data/2020_observation hours.csv')

```



## Timeline
```{r}
# add perturbation type (fitness/nonsocial/social)
fitness <- c('2020-03-18', '2020-03-19', '2020-03-20') # '2020-03-18' -> introduction, only 25 agonistic interactions observed
capture_dates <- c('2020-03-17', '2020-03-21', '2020-03-30') # intro, nonsocial (none removed), social (removal)
nonsocial <- c( '2020-03-22','2020-03-23', '2020-03-24', 
               '2020-03-27', '2020-03-28', '2020-03-29') #perturbation = 21 until 29 March
social<- c('2020-03-31','2020-04-01','2020-04-02') # perturbation = 30 March until April 2

date <- sort(unique(behave_clean$date))
bin <- c(1,1,1,2,2,2,2,3,3,3,4,4,4,4)
bin.dates <- data.frame(date, bin) 
bin.dates <- bin.dates %>%
  mutate(period=if_else(date %in% fitness,"fitness",
                        if_else(date %in% nonsocial, "nonsocial",
                                if_else(date %in% social, "social", "capture"))))


```


## Bird list

change for each perturbation as different birds are removed
```{r}
# bird list
birdIDs_capture<- read.csv("./data/2020_capture_file.csv") 

birdIDs_capture %>%
  group_by(date) %>%
  tally() # 21 birds

birdIDs <- birdIDs_capture %>% 
  filter(date=='2020-03-17') %>%
  dplyr::select(bandID, tagID)

##list of all valid color combinations

# change each time
bird.list <- sort(unique(birdIDs$tagID)) # change for social period
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

```

## check data

Include bin and period to behavior file + orderkey
```{r}
behave_bin <- behave_clean %>%
  left_join(bin.dates) %>%
  filter(period!="capture") # remove capture events as those are stressful for the birds

unique(behave_bin$bin)
head(behave_bin)
tail(behave_bin)

# how many days per period
day_sum<- behave_bin %>% 
  group_by(period, bin) %>%
  summarise(n=n_distinct(date))

## check for removed birds in social perturbation period
# GBB and BBR
check_social_actor<-filter(behave_bin, period=="social", 
                           actor == "BBR"|actor =="GBB")
check_social_subject<-filter(behave_bin, period=="social", 
                             subject == "BBR"|subject =="GBB")
check_social<-bind_rows(check_social_actor,check_social_subject)
length(behave_bin$orderkey)-length(check_social$orderkey) 

#behave_correct<-anti_join(behavekey, check_social, by = "orderkey")
#length(behave_correct$orderkey)

```


# Remove duplicates
BY bin

more conservative: removing duplicates using crowds and displacements combined from the beginning, results in less agonistic events but mistakes per behavior (entered crowd while it had to be displacement) are excluded. 

```{r remove duplicates conservative method X bin}
#### Remove duplicates for displacements and crowds ####
##more conservative: removing duplicates by crowds and displacements combined
#behave_totalagg<-behave_totalagg_removal
unique(behave_bin$behavior)

# filter displacement & crowd data from all data
aggDC <- behave_bin %>% 
  subset(behavior=="displace" | behavior=="crowd", 
         bin!=4) %>%
  dplyr::select(orderkey,  sessionKEY, period, bin, date, time, actor, subject, behavior)

str(aggDC)
length(aggDC$sessionKEY)

# keep only observations where the actor is in the list of bird IDs
aggDC.goodID <- subset(aggDC, actor %in% bird.list)

# keep only observations where the actor (found above) and the subject are both in the list of bird IDs
aggDC.goodID <- subset(aggDC.goodID, subject %in% bird.list)


## for total agonistic events (crowds and displacements combined)
#finds n CD and crowds by sessionKEY, date, time, actor, subject, behavior
dyad.totalagg.key <- aggDC.goodID %>% 
  dplyr::group_by(sessionKEY, bin, date, time,actor, subject) %>%  # removed , behavior
  tally() # counts behaviors that were observed within the same minute

#finds max agg per actor by date, time, and behavior type. Return just that summary
dyad.totalagg.maxkey <- dyad.totalagg.key %>% 
  group_by(actor, subject,bin,  date, time) %>% # removed behavior, behavior
  slice(which.max(n)) %>% # n.aggXaggXkey if summarise, n if tally
  ungroup()
#head(dyad.totalagg.maxkey)

#finds n disps per actor by date for summarized/trimmed data
dyad.totalaggXbin <- dyad.totalagg.maxkey %>% 
  group_by(actor, subject, bin) %>% # removed behavior to get total aggression
  summarise(n.aggXbin=sum(n)) %>% 
  ungroup()
#head(dyad.totalaggXbin)
sum(dyad.totalaggXbin$n.aggXbin)


#Here we are interested in total number of interactions instead of chronological #order of interactions. If you require order of interactions, then next steps will #be different from the following setup as you will have to add the behaviors with #n>1 as separate rows. 

# save summarized aggression X bin df
#write.csv(dyad.totalaggXbin, "./output/dyad.aggXbin2020.csv")

```


# Helper functions & data
```{r echo=FALSE}
# include calibri as text font
windowsFonts(Calibri = windowsFont("Calibri"))
#windowsFonts()


#make matrix function (from dataframe, 1st col with row names)
matrix.please<-function(x) {
  m<-as.matrix(x[,-1])
  rownames(m)<-x[,1]
  m
}



```


# Analyses
##Total aggression
```{r sum total aggression}
## Summarize displacements & crowds combined --> total aggression

#head(dyad.totalaggXbin) # crowds and displacement combined from the start, which is more conservative
str(dyad.totalaggXbin)
dyad.totalaggXbin <- dyad.totalaggXbin %>%
  filter(bin!=4)

# check 
unique(dyad.totalaggXbin$actor)
length(unique(dyad.totalaggXbin$actor))
unique(dyad.totalaggXbin$subject)
length(unique(dyad.totalaggXbin$subject))
sort(unique(dyad.totalaggXbin$bin))

sum(dyad.totalaggXbin$n.aggXbin)

# summarize total aggression events per bin
sum.aggDCXbin_conservative <-dyad.totalaggXbin %>%
  group_by(bin) %>% 
  filter(bin!=4) %>%
  summarize(n.agg = sum(n.aggXbin), 
            n.dyads=length(actor), 
            n.actors=n_distinct(actor),
            n.subjects=n_distinct(subject))
#head(sum.aggDCXbin_conservative)
sum.aggDCXbin_conservative

#write.csv(sum.aggDCXbin_conservative, "./output/aggression_summ2020.csv")


n.aggDC.conservative <- sum(sum.aggDCXbin_conservative$n.agg)

# check 
sum(dyad.totalaggXbin$n.aggXbin)==sum(sum.aggDCXbin_conservative$n.agg)
sort(unique(dyad.totalaggXbin$actor)) == sort(bird.list)






```
In 2020, we observed 3148 interactions (excluding unknown individuals and duplicates)



## Fig 1a) aggression rate controlled for person hours observed and group size

#### observation hours
Total hours is calculated without pee breaks as observations continued. Calculate for
  total hours per bin




```{r observation hours X bin}

bin.dates$n_observers <- c(3,3,3,3,3,3,3,3,4,4,4,4,4,4)

obs_hours <- obs.hours %>%
  separate(date, into=c("day", "month", "year")) %>%
  unite(date, year, month, day, sep = "-")

## lubridate: hm(..., quiet = FALSE, roll = FALSE)
obs_hours$start <-  lubridate::hm(obs_hours$start)
obs_hours$end <-  lubridate::hm(obs_hours$end)

# we are missing observation hours for the first 2 days (march 18 and 19). Get min and max time from behavioral dataframe
obshoursday <- behave_clean %>%
  filter(date %in% c("2020-03-18", "2020-03-19")) %>%
  group_by(date) %>%
  summarize(min=min(time),
            max=max(time)) 

obshoursday$min <-  lubridate::hm(obshoursday$min)
obshoursday$max <-  lubridate::hm(obshoursday$max)

obshoursday <- obshoursday %>%
   mutate(totaltime=lubridate::as.duration(max-min), # calculate total time
         totaltime_min=as.numeric(max-min, units="mins"),
         totaltime_hour=as.numeric(max-min, units="hours")) %>%
  summarize(sum=sum(totaltime_hour))



## total observation hours
hoursobs <- obs_hours %>%
  left_join(bin.dates, by="date") %>%
  filter(period!="capture") %>%
  mutate(totaltime=lubridate::as.duration(end-start), # calculate total time
         totaltime_min=as.numeric(end - start, units="mins"),
         totaltime_hour=as.numeric(end - start, units="hours")) %>%
  filter(name=="observations") %>%
  group_by(bin, date, n_observers) %>%
  summarize(hours_obs=sum(totaltime_hour)) 

bin1 <- obshoursday$sum+hoursobs$hours_obs[hoursobs$bin==1]
hoursobs$hours_obs[hoursobs$bin==1] <- bin1
sum(hoursobs$hours_obs)



sumhoursobs <- hoursobs %>% 
  filter(bin!=4) %>%
  group_by(bin) %>% 
  summarize(totalhours_obs=sum(hours_obs),
            meanhoursobs=mean(hours_obs),
            sdhoursobs=sd(hours_obs),
            n=n(),
            se=sdhoursobs/sqrt(n))

sum(sumhoursobs$totalhours_obs) # hours observed for bin 1-3

```
61.4 hours and 191.3 person hours across 12 days, with an average of 20.5 ± 3.6 (SD) hours of observation per three-day bin (n = 3 bins)


Person hours  (remove breaks and multiply observation hours by the number of observers)

```{r}

personhours <- obs_hours %>%
  left_join(bin.dates, by="date") %>%
  filter(period!="capture") %>%
  mutate(totaltime=lubridate::as.duration(end-start), # calculate total time
         totaltime_min=as.numeric(end - start, units="mins"),
         totaltime_hour=as.numeric(end - start, units="hours")) %>%
  #filter(name=="observations") %>%
  group_by(bin, date, subject) %>%
  summarize(breaks=sum(totaltime_hour)) %>%
  filter(subject=="break") %>%
  right_join(hoursobs) %>% 
    dplyr::select(-subject)
personhours[is.na(personhours)] <- 0

personhoursXbin <- personhours %>%
  mutate(totalhours=hours_obs*n_observers,
         personhours=totalhours-breaks) %>%
  group_by(bin) %>%
  summarize(personhours=sum(personhours)) %>%
  filter(bin!=4)

sum(personhoursXbin$personhours)
```

```{r}
## combine
obs.hours <- hoursobs %>% 
  filter(bin!=4) %>%
  group_by(bin) %>%
  summarize(hours_obs=sum(hours_obs)) %>%
  left_join(personhoursXbin) %>%
  dplyr::select(bin, hours_obs, personhours ) 
```


#### groups size and aggression rate X bin

```{r plot aggression controlled for group size and hours observed}


# total aggression
glimpse(sum.aggDCXbin_conservative)


# mean total of agonistic events observed per bin
summ_agXbin <- sum.aggDCXbin_conservative %>%
  summarize(mean_ag =mean(n.agg),
            sd_ag = sd(n.agg))

# mean and SD value
summ_ag_all <- sum.aggDCXbin_conservative %>% 
  filter(bin!=4) %>%
  mutate(birds=n.subjects)
summ_ag_all$birds[summ_ag_all$bin==2]<-21
summ_ag_all20 <- summ_ag_all %>%
  left_join(obs.hours, by="bin") %>%
  mutate(ag_bird= n.agg/birds, 
         ag_bird_obs= n.agg/birds/hours_obs, 
         ag_person_obs = n.agg/birds/personhours,
         mean = mean(ag_bird_obs),
         sd = sd(ag_bird_obs),
         n = n(),
         se=sd/sqrt(n))

# overall agression rate /hours (mean +- SE)
ave.ag.rate <- mean(summ_ag_all20$ag_bird_obs)
sd.ag.rate <- sd(summ_ag_all20$ag_bird_obs)

# overall agression rate /person hours (mean +- SE)
# ave.ag.rate <- mean(summ_ag_all$ag_person_obs)
# sd.ag.rate <- sd(summ_ag_all$ag_person_obs)


  


```

#### plot aggression rate

```{r rate = n/personhours/groupsize}
plot_aggression_control2020 <-  summ_ag_all %>%
  filter(bin!=4) %>%
  ggplot( aes(x = bin)) +
  
 # add average rate of aggression
  geom_hline(yintercept = ave.ag.rate, color = alpha("grey", 0.6), size=2) +
  
  # aggression / bird / person hours observed
  geom_line(aes(y=ag_person_obs), size=1) + 
  geom_point(aes(y=ag_person_obs), size=3) +
  
  scale_y_continuous(limits=c(0.0, 2.0)) +
  scale_x_continuous(limits = c(1, 3), # change to 4 with removal period included
                     breaks= c(1,2,3)) +
  
  theme_classic() +
  theme(legend.position="bottom",
         text = element_text(size=14),
         axis.text = element_text(size=12),
         axis.title.x = element_blank()) +
   #     axis.text.x = element_blank(),
    #    axis.ticks.x=element_line()) +
  labs(y="aggression rate")  #, title = "Amount of aggression controlled for group size and hours observed"
  
 # geom_vline(xintercept = 1.5, linetype="solid", color =  alpha("black", 0.6)) + # nonsocial perturbation
#  geom_vline(xintercept = 3.5, linetype="dashed", color =  alpha("black", 0.6)) + # removal of 2 birds (top-ranked and bottom-ranked injured bird)
  
 
 
plot_aggression_control2020
```

```{r rate = n/hours/groupsize}
plot_aggression_control2020_hours <-  summ_ag_all20 %>%
  filter(bin!=4) %>%
  ggplot( aes(x = bin)) +
  
 # add average rate of aggression
  geom_hline(yintercept = ave.ag.rate, color = alpha("grey", 0.6), size=2) +
  
  # aggression / bird / hours observed
  geom_line(aes(y=ag_bird_obs), size=1) + 
  geom_point(aes(y=ag_bird_obs), size=3) +
  
  scale_y_continuous(limits=c(0.0, 5.5),
                     breaks = c(1:5)) +
  scale_x_continuous(limits = c(1, 3), # change to 4 with removal period included
                     breaks= c(1,2,3)) +
  
  theme_classic() +
  theme(legend.position="bottom",
         text = element_text(size=14),
         axis.text = element_text(size=12),
         axis.title.x = element_blank()) +
   #     axis.text.x = element_blank(),
    #    axis.ticks.x=element_line()) +
  labs(y="aggression rate")  #, title = "Amount of aggression controlled for group size and hours observed"
  
 # geom_vline(xintercept = 1.5, linetype="solid", color =  alpha("black", 0.6)) + # nonsocial perturbation
#  geom_vline(xintercept = 3.5, linetype="dashed", color =  alpha("black", 0.6)) + # removal of 2 birds (top-ranked and bottom-ranked injured bird)
  
 
plot_aggression_control2020_hours
```


# Individual dominance rank X bin 
This is the code to identify the top-ranked bird in the bins prior the removal. 

```{r indivdidual rank}
dyad.totalaggXbin <-read.csv("./output/dyad.aggXbin2020.csv")

# check data
sort(unique(dyad.totalaggXbin$actor))
sort(unique(dyad.totalaggXbin$subject))
length(unique(dyad.totalaggXbin$actor))
length(unique(dyad.totalaggXbin$subject))


#### DOMINANCE RANK BY bin####

#head(dyad.totalaggXbin)

# make empty data frame to put data in
ec_rank <- data.frame(run.code=character(),
                      bird.list=character(), 
                      dom_ec=numeric(),
                      rank=numeric())

# run = 1

start.time <- Sys.time()

# select only bins prior removal
rank_assessment_bins <- c(1,2,3)


#i=1

for(i in seq_along(rank_assessment_bins)){
  run.code <- rank_assessment_bins[i]
  print(run.code)
  run.data <- subset(dyad.totalaggXbin, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
  data_alldyads <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 20 birds: 380 dyads, with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$date<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)
  
  ## dominance
  dom_ec <- dom_ec(mx)

  rank <- dom_ranks(dom_ec)


  # combine in dataframe
  pool <- cbind.data.frame(run.code, 
                         bird.list,
                         dom_ec,
                         rank)

  ec_rank<-rbind(ec_rank, pool )

}

end.time <- Sys.time()

#Time to run:
end.time - start.time

# check
glimpse(ec_rank)
length(unique(ec_rank$run.code))
check <- ec_rank %>% 
  group_by(run.code) %>% 
  summarize(length(bird.list))

ec_rankXbin <- ec_rank %>%
  rename(bin=run.code, id=bird.list) %>%
  mutate(power=1-dom_ec, 
         rank=22-rank) %>%
  group_by(bin) %>%
  arrange(dom_ec)

focalbirds <- ec_rankXbin %>%
  filter(bin %in% c(3), rank==1)
#focalbirds; BBG and GBB had very similar power scores, so we decided by coin toss which bird to remove (GBB)

```



# Social dominance pattern X 3-day bin
We have to run the dominance pattern analysis separately for the removal periods as we excluded one of the group members. Therefore when we are making the matrix we have to filter out the top-ranked focal bird, which is a different ID each trial. 

```{r}
#dyad.totalaggXbin <-read.csv("./output/dyad.aggXbin.csv")
dyad.totalaggXbin<- read.csv("./output/dyad.aggXbin2020.csv") %>% 
  select(-X)

glimpse(dyad.totalaggXbin)

unique(dyad.totalaggXbin$actor)
unique(dyad.totalaggXbin$subject)
length(unique(dyad.totalaggXbin$actor))
length(unique(dyad.totalaggXbin$subject))


## for bins with all 21 birds ----
bin21birds<- c(1,2,3)
dyad.totalaggXbin_all <- dyad.totalaggXbin %>%
  filter(bin %in% bin21birds)

# change bird list
bird.list <- sort(unique(birdIDs$tagID)) # change for social period
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

unique(dyad.totalaggXbin_all$actor)
unique(dyad.totalaggXbin_all$subject)
length(unique(dyad.totalaggXbin_all$actor))
length(unique(dyad.totalaggXbin_all$subject))

## removal trial 1 ----
bin19birds <- c(4)
dyad.totalaggXbinR1 <- dyad.totalaggXbin %>%
  filter(bin %in% c(4))
unique(dyad.totalaggXbinR1$bin)

removed_birds <- c("BBR", "GBB")
birdIDs_removal1 <- birdIDs %>% 
  filter(!tagID %in% removed_birds)

# change bird list
bird.list <- sort(unique(birdIDs_removal1$tagID)) # change each time
length(bird.list)
bird.list

# list of all possible dyads
dyad.list <- expand.grid(bird.list, bird.list) #head(dyad.list)
names(dyad.list) <- c("actor", "subject")
dyad.list <- subset(dyad.list, actor!=subject)
dyad.list$dyadID <- paste(dyad.list$actor, dyad.list$subject, sep="-")

unique(dyad.totalaggXbinR1$actor) # is BBB excluded?
unique(dyad.totalaggXbinR1$subject)
length(unique(dyad.totalaggXbinR1$actor))
length(unique(dyad.totalaggXbinR1$subject))




```

## observed dominance pattern

```{r Dominance pattern using crowds and displacements_ conservative, echo=FALSE}
# order date 
#head(dyad.totalaggXbin)
dyad.totalaggXbinR1<- dyad.totalaggXbin_all %>%
    arrange( bin, by_group = FALSE) 


# make empty data frame to put data in
strategies <- data.frame(run.code=character(),
                      focus=numeric(),
                      position=numeric(),
                      strategy=character())

# run = 1
set.seed(42)
start.time <- Sys.time()

loopby<-unique(dyad.totalaggXbinR1$bin)


#i=1

for(i in seq_along(loopby)){
  run.code <- loopby[i]
  print(run.code)
  run.data <- subset(dyad.totalaggXbinR1, bin==run.code)
  
# number of dyads correct?
  #length(run.data$actor)
  #check <- sum.aggXday %>% filter(date==run.code)
  #head(check)
  
  data_alldyads <- merge(dyad.list, run.data, all.x=TRUE,
                       by=c("actor", "subject")) 
  
  #head(data_alldyads)
  #length(unique(data_alldyads$dyadID)) # with 21 birds should be 420 and with 22 462 dyads
  
  data_alldyads$bin<-run.code
  
 #fill newly-merged data with 0's where no interactions
 data_alldyads[is.na(data_alldyads)] <- 0
  
   #print a check
  check <- length(data_alldyads$actor)
  print(check)

# matrix  
  mx <- reshape2::dcast(data_alldyads, actor~subject, value.var="n.aggXbin") #head(ref.behavior1.mx)
  mx[is.na(mx)] <- 0 #for linearity measure, matrix needs to be fully filled, no NAs
  mx <- matrix.please(mx)

  # ---- aggression strategy  
  # Compute focus & position
  focus <- dom_focus(mx)
  position <- dom_position(mx)
  fp <- cbind.data.frame(focus, position)
  colnames(fp) <- c("focus", "position")
  
  #Compute blur models
  blur <- dom_make_blur_data(mx)
  dom_plot_strategy(fp, blur) # plot aggression stategies
  
  #Find strategy
  strategy <- dom_categorize_strategy(data=fp, blur_data=blur)
  
  # --- combine dataframe
pool <-  cbind.data.frame(run.code, 
                         focus, 
                         position, 
                         strategy) 

strategies <- rbind(pool, strategies)

}

end.time <- Sys.time()

#Time to run:
end.time - start.time


```

```{r change each time}

#strategies_21birds <- strategies
#strategies_removal1 <- strategies

strategies2020 <- bind_rows(strategies_21birds, 
                            strategies_removal1)
strategies2020 <- strategies2020 %>%
  rename(bin=run.code) %>%
  arrange(bin) %>%
  mutate(dominance_pattern= if_else(strategy=="downward.heuristic", "downward heuristic","bullying") ) %>%
  dplyr::select(-strategy)
                                            


```




## Fig 1b) dominance patterns
strategies take long to run, so import from file

```{r plot patterns over time}
strategies <- read.csv("output/summ_agg2020.csv")%>%
  dplyr::select(bin, dominance_pattern) %>%
  filter(bin!=4) %>%
  add_row( bin=21, dominance_pattern="downward heuristic") %>%
  add_row( bin=25, dominance_pattern="close competitor") %>%
  mutate(rank = if_else(dominance_pattern=="downward heuristic", 1, 
                        ifelse(dominance_pattern=="bullying", 2, 3))) %>% 
  mutate(use = if_else(dominance_pattern=="downward heuristic", "A", 
                        ifelse(dominance_pattern=="bullying", "B", "C"))) %>% 
  mutate(observed_pattern = factor(dominance_pattern, 
            levels =c("bullying", "close competitor","downward heuristic"),
            labels =c("Bullying", "Close competitor","Downward heuristic")))


# dominance patterns long perturbation period
plot_strategies_2020 <- ggplot(strategies, 
                               aes(bin, observed_pattern)) +
  geom_point(size = 6, shape = 21, colour = "black", aes(fill = observed_pattern )) +
  labs(title = "", y="", x="social dominance pattern assessment periods") + #title = "social dominance pattern"
  scale_fill_manual(values = c( "#0000FF","#C00000" , "#7F7F7F")) +
  theme_classic() + #theme_minimal() 
  theme(#axis.title = element_text(size=12, angle = 45), #element_blank()
        #axis.text.y =element_text(size=12, angle = 25), 
        legend.position = "none",
         text = element_text(size =14),
        axis.text = element_text(size=12),
        plot.margin = unit(c(1, 1, 0.5, 1), "cm")) +
  
  scale_x_continuous(limits = c(1, 3),
                     breaks= c(1,2,3)) +
  
  # add lines to highlight removals and reintroductions
 # geom_vline(xintercept = 1.5, linetype="solid", color =  alpha("black", 0.6)) + # nonsocial perturbation
  geom_vline(xintercept = 3.5, linetype="dashed", color =  alpha("black", 0.6))  # removal


plot_strategies_2020

```

# Plot 2020 data
```{r}

  ggarrange(plot_aggression_control2020_hours , plot_strategies_2020, 
            nrow=2, ncol=1, align = "hv",
            labels = c("(a)", "(b)")
            ) # legend = "bottom", label.x = 0.15, label.y=0.85

  ggsave("./figures/plot_Fig2020hours.pdf", width= 125, height=125, units="mm")
  
  #width= 100, height=125, units="mm")
         #width= 7, height=7, units="in") # full page  width= 7, height=11, units="in") # half page width= 125, height=250, units="mm"
```

write aggression summary
```{r}
agg_summ20 <- strategies2020 %>%
  left_join(summ_ag_all) %>%
  rename(n.totalbirds=birds,
         hours.obs=sum, 
         mean_ag_rate=mean,
         sd_ag_rate=sd)
agg_summ20$focus <- round(agg_summ20$focus , digits = 2)
agg_summ20$position <- round(agg_summ20$position , digits = 2)
agg_summ20$hours.obs <- round(agg_summ20$hours.obs , digits = 2)
agg_summ20$ag_bird <- round(agg_summ20$ag_bird , digits = 2)
agg_summ20$ag_bird_obs <- round(agg_summ20$ag_bird_obs , digits = 2)

write.csv(agg_summ20, "output/summ_agg2020.csv")

```






